---
title: "Replication: Gaps in Early Warning Systems (EWSs) during  Brazil`s Largest Dengue Epidemic"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
lang: pt-BR
output:
  html_document:
    number_sections: true
    toc: true
    code_folding: hide
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
if (.Platform$OS.type == "windows") {
  Sys.setlocale("LC_CTYPE", "Portuguese_Brazil.1252")
} else {
  Sys.setlocale("LC_CTYPE", "pt_BR.UTF-8")
}

library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(ggsci)
library(here)
library(haven)
library(readr)
library(stringr)
library(officer)
library(flextable)
library(geobr)
library(sf)
library(scales)
library(RColorBrewer)
library(patchwork)
library(gtsummary)
library(CausalImpact)
library(knitr)
library(dplyr)
library(purrr)
library(gridExtra)
library(ggrepel)

```

---

# Data Preparation

## Gathering Data

```{r gather-data}
# All arbovirsuses data:
arboviruses_full <- read_csv(here("Data/cases_all_virus_2014_2024.csv")) %>% 
  mutate(
    likely_cases = tidyr::replace_na(likely_cases, 0),
    deaths = tidyr::replace_na(deaths, 0)) %>% 
  rename(calendar_year = year,
         calendar_week = week)  %>%
  mutate(virus = str_to_title(virus),
          calendar_year_week = paste0(calendar_year, "w", calendar_week)) 

# Epidemiological week starting on the first week of the winter
epiweek40 <- read_dta(
  here("Data/epidemiological_week_startsweek40.dta")) %>% 
  rename(calendar_year_week = wy,
         epi_week = epidemiologic_week_startsweek40,
         epi_year = year_week40) %>% 
  mutate(epi_week = as.numeric(epi_week),
         epi_year = as.numeric(epi_year)) %>% 
  mutate(epi_year_week = paste0(epi_year, "w", epi_week))

#List of municipalities of interest
codibge <- list(
  "Rio Branco"      = "120040",
  "Maceió"          = "270430",
  "Macapá"          = "160030",
  "Manaus"          = "130260",
  "Salvador"        = "292740",
  "Fortaleza"       = "230440",
  "Brasília"        = "530010",
  "Vitória"         = "320530",
  "Goiânia"         = "520870",
  "São Luís"        = "211130",
  "Cuiabá"          = "510340",
  "Campo Grande"    = "500270",
  "Belo Horizonte"  = "310620",
  "Belém"           = "150140",
  "João Pessoa"     = "250750",
  "Curitiba"        = "410690",
  "Recife"          = "261160",
  "Teresina"        = "221100",
  "Rio de Janeiro"  = "330455",
  "Natal"           = "240810",
  "Porto Alegre"    = "431490",
  "Porto Velho"     = "110020",
  "Boa Vista"       = "140010",
  "Florianópolis"   = "420540",
  "São Paulo"       = "355030",
  "Aracaju"         = "280030",
  "Palmas"          = "172100")

# Get Geospatial Data
states_sf <- read_state(year = 2020)
capitals_points <- read_municipality(year = 2020) %>%
  st_centroid() %>%
  mutate(code_muni = as.double(substr(as.character(code_muni), 1, 6)))
```

---

## Preparing Data

```{r preparing-data}
# 1. Create the base dataset
arboviruses_base <- arboviruses_full %>%
  select(ibge, calendar_year, calendar_week, calendar_year_week, municipality, uf, virus, likely_cases, deaths, population) %>%
  left_join(epiweek40, by = "calendar_year_week") %>%
  arrange(epi_year, epi_week) %>%
  filter(!is.na(municipality), !is.na(epi_year), !is.na(epi_week))

# 2. Calculate population for each epidemiological year
# States
state_population <- arboviruses_base %>%
  distinct(uf, ibge, epi_year, .keep_all = TRUE) %>%
  group_by(uf, epi_year) %>%
  summarise(state_pop = sum(population), .groups = "drop")

# Country
country_population <- arboviruses_base %>%
  distinct(ibge, epi_year, .keep_all = TRUE) %>%
  group_by(epi_year) %>%
  summarise(country_pop = sum(population), .groups = "drop")

# 3. Create datasets for each geographic level

# Munis 
arboviruses_muni <- arboviruses_base %>%
  filter(ibge %in% unlist(codibge)) %>%
  group_by(ibge, municipality, uf, calendar_year, calendar_week, calendar_year_week, epi_year, epi_week, epi_year_week, virus) %>%
  summarise(
    cases = sum(likely_cases),
    deaths = sum(deaths),
    population = first(population), # pop is measured on year basis
    .groups = "drop"
  ) %>%
  mutate(
    incidence = (cases / population) * 100000,
    death_rate = (deaths / population) * 100000
  ) %>%
  # Calculate cumulative per epi year
  group_by(ibge, municipality, uf, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

arboviruses_muni <- arboviruses_muni %>%
mutate(muni_name = paste0(stringr::str_remove(municipality, "^\\d+\\s"), " (", uf, ")")) %>%
  arrange(epi_year, epi_week) %>%
  mutate(epi_year_week = factor(epi_year_week, levels = unique(epi_year_week), ordered = TRUE))

# States 
arboviruses_state <- arboviruses_base %>%
  group_by(uf, calendar_year, calendar_week, calendar_year_week, epi_year, epi_week, epi_year_week, virus) %>%
  summarise(
    cases = sum(likely_cases),
    deaths = sum(deaths),
    .groups = "drop"
  ) %>%
  full_join(state_population, by = c("uf", "epi_year")) %>%
  mutate(
    incidence = (cases / state_pop) * 100000,
    death_rate = (deaths / state_pop) * 100000
  ) %>%
  # Calculate cumulative per epi year
  group_by(uf, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

arboviruses_state <- arboviruses_state %>%
  arrange(epi_year, epi_week) %>%
  mutate(epi_year_week = factor(epi_year_week, levels = unique(epi_year_week), ordered = TRUE))

# Country 
arboviruses_country <- arboviruses_base %>%
  group_by(calendar_year, calendar_week, calendar_year_week, epi_year, epi_week, epi_year_week, virus) %>%
  summarise(
    cases = sum(likely_cases),
    deaths = sum(deaths),
    .groups = "drop"
  ) %>%
  left_join(country_population, by = "epi_year") %>%
  mutate(
    incidence = (cases / country_pop) * 100000,
    death_rate = (deaths / country_pop) * 100000,
    country = "Brazil"
  ) %>%
  # Calculate cumulative per epi year
  group_by(country, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

arboviruses_country <- arboviruses_country %>%
  arrange(epi_year, epi_week) %>%
  mutate(epi_year_week = factor(epi_year_week, levels = unique(epi_year_week), ordered = TRUE))
```

---

# Data Visualization

## Line Plots

```{r funct-plot}
create_arbovirus_plot <- function(data, y_var, virus_filter_expr, y_lab,
                                  grouping_var, legend_position,
                                  facet_by = NULL,
                                  year_interval = 1,
                                  file_name) {

  # Use quosures for tidy evaluation
  y_var_quo <- enquo(y_var)
  grouping_var_quo <- enquo(grouping_var)
  virus_filter_quo <- enquo(virus_filter_expr)

  # Find the data for the first week of each epidemiological year
  axis_break_data <- data %>%
    group_by(epi_year) %>%
    slice_head(n = 1) %>%
    ungroup() %>%
    slice(seq(1, n(), by = year_interval))
    
  # Get the epi_year_week values for the axis break positions
  axis_breaks <- axis_break_data %>% pull(epi_year_week)
  
  # Get the corresponding calendar_year_week values for the axis labels
  axis_labels <- axis_break_data %>% pull(calendar_year_week)

  # Build the plot
  p <- data %>%
    arrange(epi_year, epi_week) %>%
    filter(!!virus_filter_quo) %>%
    ggplot(aes(x = epi_year_week, y = !!y_var_quo, group = !!grouping_var_quo, color = virus)) +
    geom_line(alpha = 0.8, linewidth = 0.9) +
    scale_color_lancet(name = "Virus") +
    labs(
      x = "Epidemiological Year-Week",
      y = y_lab
    ) +
    scale_x_discrete(breaks = axis_breaks, labels = axis_labels) +
    theme_classic(base_size = 14) + 
    theme(
      text = element_text(size = 20),
      axis.title = element_text(face = "bold", size = 20),
      axis.text = element_text(size = 20),
      legend.position = legend_position,
      panel.grid.major.y = element_line(color = "grey85", linewidth = 0.5),
      axis.text.x = element_text(angle = 45, hjust = 1))

  # Add faceting if specified
  if (!is.null(facet_by)) {
    p <- p + facet_wrap(vars(!!sym(facet_by)), scales = "free_y")
  }

  # Print and save
  print(p)
  ggsave(
    here("Graphs", file_name),
    p,
    height = 15,
    width = 15,
    dpi = 600,
    create.dir = TRUE
  )
}
```

### Incidence (Line)

#### Country-wide Incidence (Line)

```{r incidence-country-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_country,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "none",
  file_name = "Incidence of Dengue in Brazil.png"
)
```

#### State-level Incidence (Line)

```{r incidence-state-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_state,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "none",
  facet_by = "uf",
  year_interval = 2,
  file_name = "Incidence of Dengue per State.png"
)
```

#### Municipalities Incidence (Line)

```{r incidence-muni-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_muni,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "none",
  facet_by = "muni_name",
  year_interval = 2,
  file_name = "Incidence of Dengue per Selected Municipalities.png"
)
```

### Mortality (Line)

#### Country-wide Deaths (Line)

```{r deaths-country-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_country,
  y_var = death_rate,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Death Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "none",
  file_name = "Deaths of Dengue in Brazil.png"
)
```

#### State-level Deaths (Line)

```{r deaths-state-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_state,
  y_var = death_rate,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Death Rate (per 100,000 pop.) per Epidemiological Year",
  grouping_var = virus,
  legend_position = "none",
  facet_by = "uf",
  year_interval = 2,
  file_name = "Deaths of Dengue per State.png"
)
```

#### Municipalities Deaths (Line)

```{r deaths-muni-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_muni,
  y_var = death_rate,
 virus_filter_expr = virus == "Dengue",
  y_lab = "Death Rate (per 100,000 pop.) per Epidemiological Year",
  grouping_var = virus,
  legend_position = "none",
  facet_by = "muni_name",
  year_interval = 2,
  file_name = "Deaths of Dengue per Selected Municipalities.png"
)
```


## Overlaid Years Plotting Function

```{r funct-plot-overlaid}
create_arbovirus_plot_overlaid <- function(data, y_var, virus_filter_expr, y_lab,
                                           facet_by = NULL,
                                           file_name,
                                           legend_position = "top") {

  # Use quosures for tidy evaluation
  y_var_quo <- enquo(y_var)
  virus_filter_quo <- enquo(virus_filter_expr)

  # Define epi-week order (winter year: 40 → 1 → 39)
  winter_week_order <- c(40:53, 1:39)

  plot_data <- data %>%
    filter(epi_year >= 2015 & epi_year <= 2024) %>%
    mutate(
      epi_year = as.factor(epi_year),
      calendar_week = factor(calendar_week, levels = winter_week_order)
    )

  # Define custom axis breaks and labels
  axis_breaks <- c(40, 1, 20, 39)
  axis_labels <- c("40", "1", "20", "39")

  # Build plot
  p <- plot_data %>%
    filter(!!virus_filter_quo) %>%
    ggplot(aes(x = calendar_week, y = !!y_var_quo, group = epi_year, color = epi_year)) +
    geom_line(alpha = 0.8, linewidth = 1.2, na.rm = TRUE) +
    scale_color_viridis_d(name = "Epidemiological Year",
                          direction = -1) +
    labs(
      x = "Calendar Week (Week 40 of Previous Calendar Year to Week 39 of Current Calendar Year)",
      y = y_lab
    ) +
    # Ensure all epi weeks appear (even if empty)
    scale_x_discrete(breaks = as.character(axis_breaks), labels = axis_labels, drop = FALSE) +
    theme_classic(base_size = 14) +
    theme(
      text = element_text(size = 20),
      axis.title = element_text(face = "bold", size = 20),
      axis.text = element_text(size = 20),
      legend.position = legend_position,
      panel.grid.major.y = element_line(color = "grey85", linewidth = 0.5),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )

  # Add faceting if requested
  if (!is.null(facet_by)) {
    p <- p + facet_wrap(vars(!!sym(facet_by)), scales = "free_y")
  }

  # Output
  print(p)
  ggsave(
    here("Graphs", file_name),
    p,
    height = 15,
    width = 15,
    dpi = 600,
    create.dir = TRUE
  )
}
```

### Incidence (Overlaid)

#### Country-wide Incidence (Overlaid)

```{r incidence-country-dengue-overlaid, fig.width=15, fig.height=15}
create_arbovirus_plot_overlaid(
  data = arboviruses_country,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  legend_position = "top",
  file_name = "Incidence of Dengue in Brazil_overlaid.png"
)
```

#### State-level Incidence (Overlaid)

```{r incidence-state-dengue-overlaid, fig.width=15, fig.height=15}
create_arbovirus_plot_overlaid(
  data = arboviruses_state,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  legend_position = "top",
  facet_by = "uf",
  file_name = "Incidence of Dengue per State_overlaid.png"
)
```

#### Municipalities Incidence (Overlaid)

```{r incidence-muni-dengue-overlaid, fig.width=15, fig.height=15}
create_arbovirus_plot_overlaid(
  data = arboviruses_muni,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  legend_position = "top",
  facet_by = "muni_name",
  file_name = "Incidence of Dengue per Selected Municipalities_overlaid.png"
)
```

### Mortality (Overlaid)

#### Country-wide Deaths (Overlaid)

```{r deaths-country-dengue-overlaid, fig.width=15, fig.height=15}
create_arbovirus_plot_overlaid(
  data = arboviruses_country,
  y_var = death_rate,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Death Rate (per 100,000 pop.)",
  legend_position = "top",
  file_name = "Deaths of Dengue in Brazil_overlaid.png"
)
```

#### State-level Deaths (Overlaid)

```{r deaths-state-dengue-overlaid, fig.width=15, fig.height=15}
create_arbovirus_plot_overlaid(
  data = arboviruses_state,
  y_var = death_rate,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Cumulative Death Rate (per 100,000 pop.) per Epidemiological Year",
  legend_position = "top",
  facet_by = "uf",
  file_name = "Deaths of Dengue per State_overlaid.png"
)
```

#### Municipalities Deaths (Overlaid)

```{r deaths-muni-dengue-overlaid, fig.width=15, fig.height=15}
create_arbovirus_plot_overlaid(
  data = arboviruses_muni,
  y_var = death_rate,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Cumulative Death Rate (per 100,000 pop.) per Epidemiological Year",
  legend_position = "top",
  facet_by = "muni_name",
  file_name = "Deaths of Dengue per Selected Municipalities_overlaid.png"
)
```

## Mapping Incidence and Mortality in 2024

```{r mapping-dengue, fig.width=16, fig.height=12}
# Define the Epidemiological Year of Interest
epi_year_to_plot <- 2024

# Combine Data with Map
dengue_states_map_data <- arboviruses_state %>%
  filter(epi_year == epi_year_to_plot,
         virus == "Dengue") %>%
  slice_max(order_by = epi_week, n = 1) %>%
  left_join(states_sf, ., by = c("abbrev_state" = "uf"))

dengue_capitals_map_data <- arboviruses_muni %>%
  filter(epi_year == epi_year_to_plot,
         virus == "Dengue",
         ibge %in% codibge) %>%
  slice_max(order_by = epi_week, n = 1) %>%
  left_join(capitals_points, ., by = c("code_muni" = "ibge")) %>% 
  filter(!is.na(municipality), !is.na(epi_year), !is.na(epi_week))

# Define breaks for Incidence
inc_breaks <- c(0, 500, 2500, Inf) 
inc_labels <- c("0 - 499", "500 - 2,499", "2,500+")

# Define breaks for Death Rate
dengue_death_rate_breaks <- c(0, 2, 5, Inf) 
dengue_death_rate_labels <- c("0 - 1.9", "2 - 4.9", "5+")

# Apply these unified breaks and clean labels to both datasets
dengue_states_map_data <- dengue_states_map_data %>%
  mutate(
    incidence_cat = cut(cumulative_incidence, breaks = inc_breaks, labels = inc_labels, include.lowest = TRUE, right = FALSE),
    death_rate_cat = cut(cumulative_death_rate, breaks = dengue_death_rate_breaks, labels = dengue_death_rate_labels, include.lowest = TRUE, right = FALSE)
  )
dengue_capitals_map_data <- dengue_capitals_map_data %>%
  mutate(
    incidence_cat = cut(cumulative_incidence, breaks = inc_breaks, labels = inc_labels, include.lowest = TRUE, right = FALSE),
    death_rate_cat = cut(cumulative_death_rate, breaks = dengue_death_rate_breaks, labels = dengue_death_rate_labels, include.lowest = TRUE, right = FALSE)
  )

# Generate the Maps

### MAP 1: CUMULATIVE INCIDENCE

dengue_incidence_map_themed <- ggplot() +
  geom_sf(data = dengue_states_map_data, aes(fill = incidence_cat), color = "black", linewidth = 0.5) +
  geom_sf(data = dengue_capitals_map_data, aes(size = incidence_cat), shape = 21, fill = NA, color = "darkgrey", alpha = 1, stroke = 2) +
   scale_fill_brewer(
    palette = "Greens",
    name = "Cumulative Incidence (States)\n(per 100,000 pop.)",
    na.translate = FALSE,
    drop = TRUE 
  ) +
  scale_size_manual(
    name = "Cumulative Incidence (Capitals)\n(per 100,000 pop.)",
    values = c(3, 6, 9),
    drop = FALSE
  ) +
   guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2)
  ) +
  theme_bw() +
  theme(
    legend.position = c(0.8, 0.1),
    legend.justification = c("left", "bottom"),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.title = element_text(),
    axis.text = element_text(size=12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16))

print(dengue_incidence_map_themed)
ggsave(here("Graphs", paste0("Map Dengue Incidence 2024.png")), dengue_incidence_map_themed, width = 16, height = 12, dpi = 600)
  
### MAP 2: CUMULATIVE DEATH RATE

dengue_deaths_rate_map_themed <- ggplot() +
  geom_sf(data = dengue_states_map_data, aes(fill = death_rate_cat), color = "black", linewidth = 0.5) +
  geom_sf(data = dengue_capitals_map_data, aes(size = death_rate_cat), 
           shape = 21, fill = NA, color = "darkgrey", alpha = 1,stroke = 2) +
  scale_fill_brewer(
    palette = "Purples",
    name = "Cumulative Death Rate (States)\n(per 100,000 pop.)",
    na.translate = FALSE,
    drop = FALSE
  ) +
  scale_size_manual(
    name = "Cumulative Death Rate (Capitals)\n(per 100,000 pop.)",
    values = c(3, 6, 9),
    drop = FALSE
  ) +
   guides(
    fill = guide_legend(order = 1),
    size = guide_legend(order = 2)
  ) +
  theme_bw() +
  theme(
    legend.position = c(0.8, 0.1),
    legend.justification = c("left", "bottom"),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.title = element_text(),
    axis.text = element_text(size=12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 16))

print(dengue_deaths_rate_map_themed)
ggsave(here("Graphs", paste0("Map Dengue Death 2024.png")), dengue_deaths_rate_map_themed, width = 16, height = 12, dpi = 600)
```

---

# Risk Assessment

## Risk assessment function

```{r risk-function}
risk_analysis <- function(df, analysis_year = 2024) {
  
  # === PART 1: PREPARE DATA & BUILD ENDEMIC CHANNEL ===
  
  # Create the endemic channel using all years >= 2015, excluding the analysis year
  endemic_channel <- df %>%
    filter(epi_year >= 2015 & epi_year != analysis_year) %>%
    group_by(epi_week) %>%
    summarise(
      median_incidence = median(incidence),
      high = quantile(incidence, 0.75),
      low = quantile(incidence, 0.25),
      .groups = "drop"
    )
  
  # Isolate the current year's data for the analysis
  current_year_data <- df %>%
    filter(epi_year == analysis_year) %>%
    select(calendar_year_week, epi_week, incidence_current = incidence, deaths_current = deaths)
  
  # Combine endemic channel and current year data
  
  max_week <- max(df$epi_week[df$epi_year == analysis_year], na.rm = TRUE) # Dynamically find the maximum epi_week for the specified analysis_year
  
  risk_final <- tibble(epi_week = 1:max_week) %>%  
    left_join(endemic_channel, by = "epi_week") %>%
    left_join(current_year_data, by = "epi_week") %>%
    # Replace NA values with 0 to allow for calculations
    mutate(
      across(everything(), ~replace_na(., 0))
    )
  
  # === PART 2: APPLY RISK ASSESSMENT RULES ===
  
  risk_final <- risk_final %>%
    arrange(epi_week) %>%
    mutate(
      # Condition 1: Increased incidence between median and high threshold for 4 weeks
      increased_incidence_within = ifelse(incidence_current > median_incidence & incidence_current < high , 1, 0),
      increased_incidence_within_4w = zoo::rollapply(
        increased_incidence_within, 4,
        function(x) as.integer(all(x == 1)),
        fill = 0, align = "right"),
      
      # Condition 2: Increased incidence above high threshold for 4 weeks
      increased_incidence_high = ifelse(incidence_current > high , 1, 0),
      increased_incidence_high_4w = zoo::rollapply(
        increased_incidence_high, 4,
        function(x) as.integer(all(x == 1)),
        fill = 0, align = "right"),
      
      # Condition 3: Increased deaths for 4 weeks
      increased_deaths = ifelse(deaths_current > lag(deaths_current, 1) , 1, 0),
      increased_deaths_4w = zoo::rollapply(
        increased_deaths, 4,
        function(x) as.integer(all(x == 1)),
        fill = 0, align = "right"),
      
       # Condition 4: Deaths per cases above 0.05%
      lethality = ifelse(incidence_current > 0, (deaths_current / incidence_current) > 0.0005, 0),
      
      # Condition 5: Exponential increase in incidence AND incidence above high threshold 
      exponential_incidence = ifelse(incidence_current > 0 & lag(incidence_current, 1) > 0,
                                     log(incidence_current) > log(lag(incidence_current, 1)) &
                                       incidence_current > high, 0),
      
      # Condition 6: Increased deaths for more than 4 weeks (5 weeks)
      increased_deaths_5w = zoo::rollapply(
        increased_deaths, 5,
        function(x) as.integer(all(x == 1)),
        fill = 0, align = "right"),
        
    # Assign a risk level based on a nested, hierarchical logic
      riskassessment = case_when(
      epi_week <= 4 ~ -1L,
      
      # EPIDEMIC - Condition 5 AND/OR Condition 6
      (exponential_incidence == 1 & increased_deaths_5w == 1) |
      (exponential_incidence == 1 | increased_deaths_5w == 1) ~ 3L,
      
      # ALERT - Condition 2 AND/OR Condition 3 AND/OR Condition 4
      (increased_incidence_high_4w == 1 & increased_deaths_4w == 1 & lethality == 1) |
      (increased_incidence_high_4w == 1 & increased_deaths_4w == 1) |
     (increased_incidence_high_4w == 1 & lethality == 1) |
      (increased_deaths_4w == 1 & lethality == 1) |
     (increased_incidence_high_4w == 1 | increased_deaths_4w == 1 | lethality == 1) ~ 2L,
  
      # Mobilization - Condition 1
      increased_incidence_within_4w == 1 ~ 1L,
   
        # Default: Normality
        TRUE ~ 0L
      )
    ) 
  
    return(list(risk_final = risk_final))
}
```

### Running Risk assessment function

```{r risk-func-run}
# Country Level
nested_data_country <- arboviruses_country %>%
  group_by(country, virus) %>%
  nest()

results_country <- nested_data_country %>%
  mutate(risk_analysis_results = map(data, risk_analysis))

results_country_filtered <- results_country %>%
  filter(!map_lgl(risk_analysis_results, is.null))

risk_2024_country <- results_country_filtered %>%
  mutate(risk_data = map(risk_analysis_results, ~ .x$risk_final)) %>%
  select(country, virus, risk_data) %>%
  unnest(cols = c(risk_data))

# State Level
nested_data_state <- arboviruses_state %>%
  group_by(uf, virus) %>%
  nest()

results_state <- nested_data_state %>%
  mutate(risk_analysis_results = map(data, risk_analysis))

results_state_filtered <- results_state %>%
  filter(!map_lgl(risk_analysis_results, is.null))
  
risk_2024_state <- results_state_filtered %>%
  mutate(risk_data = map(risk_analysis_results, ~ .x$risk_final)) %>%
  select(uf, virus, risk_data) %>%
  unnest(cols = c(risk_data))

# Municipality Level
nested_data_muni <- arboviruses_muni %>%
  group_by(municipality, muni_name, virus, ibge) %>%
  nest()

results_muni <- nested_data_muni %>%
  mutate(risk_analysis_results = map(data, risk_analysis))

results_muni_filtered <- results_muni %>%
  filter(!map_lgl(risk_analysis_results, is.null))

risk_2024_muni <- results_muni_filtered %>%
  mutate(risk_data = map(risk_analysis_results, ~ .x$risk_final)) %>%
  select(municipality, muni_name, ibge, virus, risk_data) %>%
  unnest(cols = c(risk_data))
```

### Summary Table of Risk Assessment

```{r table-risk-summary-unified}
# --- Create a unified location column for each level ---

# Country
risk_country_prep <- risk_2024_country %>%
  filter(virus == "Dengue") %>%
  mutate(location = country, level = "Country") %>%
  select(level, location, epi_week, calendar_year_week, riskassessment)

# State
risk_state_prep <- risk_2024_state %>%
  filter(virus == "Dengue") %>%
  mutate(location = uf, level = "State") %>%
  select(level, location, epi_week, calendar_year_week, riskassessment)

# Municipality
risk_muni_prep <- risk_2024_muni %>%
  filter(virus == "Dengue") %>%
  mutate(location = muni_name, level = "Municipality") %>%
  select(level, location, ibge, epi_week, calendar_year_week, riskassessment)

# --- Combine all levels into one dataframe ---
all_levels_risk <- bind_rows(risk_country_prep, risk_state_prep, risk_muni_prep)

# --- Calculate metrics ---
summary_data <- all_levels_risk %>%
  group_by(level, location) %>%
  summarise(
    # Find the first epidemiological week for alert (risk level 2)
    first_alert_week_epi = min(epi_week[riskassessment == 2], na.rm = TRUE),
    
    # Find the first epidemiological week for epidemic (risk level 3)
    first_epidemic_week_epi = min(epi_week[riskassessment == 3], na.rm = TRUE),

    # Calculate max consecutive weeks at level 2 using run-length encoding (rle)
    max_consecutive_level2 = {
      rle_res <- rle(riskassessment == 2)
      if (any(rle_res$values)) max(rle_res$lengths[rle_res$values]) else 0
    },

    # Calculate max consecutive weeks at level 3
    max_consecutive_level3 = {
      rle_res <- rle(riskassessment == 3)
      if (any(rle_res$values)) max(rle_res$lengths[rle_res$values]) else 0
    },

    # Calculate max consecutive weeks at level 2 or 3
    max_consecutive_level2_or_3 = {
      rle_res <- rle(riskassessment %in% c(2, 3))
      if (any(rle_res$values)) max(rle_res$lengths[rle_res$values]) else 0
    },
    
    .groups = 'drop'
  )

# --- Prepare for final table creation ---
# Create a distinct lookup table for week information to prevent join issues
week_lookup <- all_levels_risk %>% 
  select(location, epi_week, calendar_year_week) %>% 
  distinct()

# Join to get calendar weeks and format the final table
summary_table <- summary_data %>%
  # Join to get the calendar week for the first alert
  left_join(week_lookup, by = c("location", "first_alert_week_epi" = "epi_week")) %>%
  rename(first_alert_cal = calendar_year_week) %>%
  # Join to get the calendar week for the first epidemic
  left_join(week_lookup, by = c("location", "first_epidemic_week_epi" = "epi_week")) %>%
  rename(first_epidemic_cal = calendar_year_week) %>%
  # Handling cases with no alert/epidemic
  mutate(
    first_alert_display = if_else(is.infinite(first_alert_week_epi), "No Alert", first_alert_cal),
    first_epidemic_display = if_else(is.infinite(first_epidemic_week_epi), "No Epidemic", first_epidemic_cal)
  ) %>%
  # Select and rename columns for the final output
  select(
    `Level` = level,
    `Location` = location,
    `First Week Alert (level 2)` = first_alert_display,
    `Max. Consecutive Weeks (level 2)` = max_consecutive_level2,
    `First Week Epidemic (level 3)` = first_epidemic_display,
    `Max. Consecutive Weeks (level 3)` = max_consecutive_level3,
    `Max. Consecutive Weeks (level 2 or 3)` = max_consecutive_level2_or_3
  )

# --- Display table ---
knitr::kable(summary_table, caption = "Risk Assessment Summary for All Levels (Dengue, 2024)")

# --- Save to a Word document ---
doc <- read_docx() %>%
  body_add_par("Risk Assessment Summary for All Levels (Dengue, 2024)", style = "heading 1") %>%
  body_add_flextable(flextable(summary_table))

print(doc, target = "risk_assessment.docx")
```

#### Combining Risk Data with Emergency Declarations in Epidemiological Year 2024

```{r preparing-risk-data}
# Country Risk Assessment
risk_2024_country <- risk_2024_country %>%
  mutate(epi_week = as.numeric(epi_week),
         riskassessment = as.factor(riskassessment))

# States Risk Assessment
risk_2024_state <- risk_2024_state %>%
  mutate(epi_week = as.numeric(epi_week),
         riskassessment = as.factor(riskassessment))

# Municipalities Risk Assessment
risk_2024_muni <- risk_2024_muni %>%
  mutate(epi_week = as.numeric(epi_week),
         riskassessment = as.factor(riskassessment))

# Creation of the Centro de Operações de Emergências de Saúde Pública para Dengue e outras Arboviroses, no âmbito do Ministério da Saúde
emergency_country <- tibble::tibble(
  label = "02/02/2024",
  date_issued = dmy("02/02/2024")
) %>%
  mutate(
    year = year(date_issued),
    week = week(date_issued), # Get standard week number
    calendar_year_week = paste0(year, "w", as.integer(week)) # Create a join key (e.g., "2024w5")
  ) %>%
  left_join(epiweek40, by = "calendar_year_week") 

# State Emergency Declarations
emergency_state <- tibble::tribble(
    ~uf, ~date_issued,
    "AC", "05/01/2024",
    "DF", "25/01/2024",
    "GO", "02/02/2024",
    "ES", "21/02/2024",
    "RJ", "21/02/2024",
    "SC", "22/02/2024", 
    "MG", "26/02/2024",
    "AP", "28/02/2024",
    "SP", "05/03/2024",
    "RS", "12/03/2024",
    "PR", "14/03/2024"
  ) %>%
  mutate(
    date_issued = dmy(date_issued),
    year = year(date_issued),
    week = week(date_issued), # Get standard week number
    calendar_year_week = paste0(year, "w", as.integer(week)) # Create a join key
  ) %>%
  left_join(epiweek40, by = "calendar_year_week") 

# Municipal Emergency Declarations
emergency_muni <- tibble::tribble(
    ~ibge, ~date_issued,
    420540, "29/11/2023",  # Florianópolis
    330455, "02/02/2024",  # Rio de Janeiro
    310620, "16/02/2024",  # Belo Horizonte
  #  420540, "22/02/2024",  # Florianópolis 
    240810, "01/03/2024",  # Natal
    520870, "13/03/2024",  # Goiânia
    355030, "18/03/2024",  # São Paulo
    431490, "22/04/2024"   # Porto Alegre
  ) %>%
  mutate(
    ibge = as.numeric(ibge), # Ensure type matches other data
    date_issued = dmy(date_issued),
    year = year(date_issued),
    week = week(date_issued),
    calendar_year_week = paste0(year, "w", as.integer(week))
  ) %>%
  left_join(epiweek40, by = "calendar_year_week") %>%
  left_join(
    arboviruses_muni %>% distinct(ibge, muni_name),
    by = "ibge"
  )

# Aggregate labels for states declaring emergency in the same week
emergency_state_labels <- emergency_state %>%
  group_by(epi_week) %>%
  summarise(uf_labels = paste(uf, collapse = "\n"))

# Create the state emergency data needed for the municipal plots
muni_info <- arboviruses_muni %>%
  distinct(muni_name, uf)

emergency_state_for_muni_plot <- emergency_state %>%
  left_join(muni_info, by = "uf", relationship = "many-to-many") %>%
  filter(!is.na(epi_week))
```

## Visualizing Risk Assessment and Emergency Declarations in 2024

```{r function-risk-plot}
create_risk_plot <- function(data, facet_var = NULL) {
  
  # Define axis breaks and find their corresponding calendar week labels
  axis_breaks <- seq(min(data$epi_week), max(data$epi_week), by = 13) %>% 
    union(max(data$epi_week))
  axis_labels <- data$calendar_year_week[match(axis_breaks, data$epi_week)]
  
  # Order risk assessment as factor
  all_levels <- c("-1", "0", "1", "2", "3")
  
  plot_data <- data %>%
    mutate(riskassessment = factor(riskassessment, levels = all_levels))
  
  dummy_data <- tibble(riskassessment = factor(all_levels, levels = all_levels))
  
  # Base plot
  p <- plot_data %>%
    ggplot(aes(x = epi_week)) +
    # Incidence bars colored by risk level
    geom_col(aes(y = incidence_current, fill = riskassessment), width = 1, alpha = 0.8) +
    # Endemic channel 
    geom_ribbon(aes(ymin = low, ymax = high, fill = "Endemic Channel", group = 1), alpha = 0.8) +
    # Median incidence line
    geom_line(aes(y = median_incidence, color = "Historical Median", linetype = "Historical Median", group = 1), linewidth = 0.8) +
    # Add an invisible layer that uses the dummy data to make sure all factors are present
    geom_tile(data = dummy_data, mapping = aes(x = -Inf, y = -Inf, fill = riskassessment), alpha = 0) +
    # Scale x-axis
    scale_x_continuous(breaks = axis_breaks, labels = axis_labels) +
    # Define custom colors and labels for the fill aesthetic (Risk Levels)
    scale_fill_manual(
       name = "Risk Level & Baseline",
  values = c(
    "Endemic Channel" = "darkgrey",
    "-1" = "lightgrey",
    "0" = "green",
    "1" = "yellow",
    "2" = "orange",
    "3" = "red"),
  labels = c(
    "Endemic Channel" = "Endemic Channel",
    "-1" = "Learning",
    "0" = "Level 0 - Normality",
    "1" = "Level 1 - Mobilization",
    "2" = "Level 2 - Alert",
    "3" = "Level 3 - Epidemic"), 
  drop = FALSE,
  na.translate = T,
  limits = c("Endemic Channel", "-1", "0", "1", "2", "3")) +
  # Define manual color for historical Median line
  scale_color_manual(
  values = c("Historical Median" = "blue")) +
  # Define manual linetype for historical Median line
  scale_linetype_manual(
  values = c("Historical Median" = "dashed")) +
  guides(
      fill = guide_legend(
        title = NULL,
        order = 1,
        nrow = 1, 
        override.aes = list(alpha = 1) # 
      ),
      color = guide_legend(
        title = NULL,
        order = 2,
        override.aes = list(linetype = "dashed", color = "blue")
      ),
      linetype = "none"
    ) +
theme_classic(base_size = 14) +
    theme(
      text = element_text(size = 18),
      axis.title = element_text(face = "bold", size = 18),
      axis.text = element_text(size = 18),
      legend.position = "bottom",
      legend.title = element_text(face = "bold", size = 18),
      legend.text = element_text(size = 18),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.5), # Add subtle grid lines
      strip.background = element_rect(fill = "grey92", color = NA), # Style facet titles
      strip.text = element_text(face = "bold", size = rel(0.9)),
      legend.box = "vertical",
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    labs(
      y = "Incidence Rate (per 100,000)", x = "Epidemiological Week")

  # Add faceting if a facet variable is provided
  if (!is.null(facet_var)) {
    p <- p + facet_wrap(vars(!!sym(facet_var)), scales = "free_y")
  }
  
  return(p)
}
```

### Country

```{r risk-country-dengue, fig.width=16, fig.height=12}
dengue_risk_country <- create_risk_plot(
  data = risk_2024_country %>% filter(virus == "Dengue")) +
  geom_vline(data = emergency_country, aes(xintercept = epi_week), linetype = "solid", color = "black", size = 1) +
  geom_vline(data = emergency_state, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1) +
  geom_text(data = emergency_state_labels, aes(x = epi_week, y = 220, label = uf_labels), vjust = 1.5, size = 6)

print(dengue_risk_country)
ggsave(here("Graphs/Risk Levels of Dengue Outbreak in 2024 in Brazil with Emergency Dec.png"), dengue_risk_country, height = 12, width = 16, dpi = 600, create.dir = TRUE)
```

### States 

```{r risk-state-dengue, fig.width=16, fig.height=12}
dengue_risk_state <- create_risk_plot(
  data = risk_2024_state %>% filter(virus == "Dengue"),
  facet_var = "uf") + 
  geom_vline(data = emergency_state, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1.5)

print(dengue_risk_state)
ggsave(here("Graphs/Risk Levels of Dengue Outbreak in 2024 per State with Emergency Dec.png"), dengue_risk_state, height = 12, width = 16, dpi = 600, create.dir = TRUE)
```
    
### Municipalities

```{r risk-muni-dengue, fig.width=16, fig.height=12}
dengue_risk_muni <- create_risk_plot(
  data = risk_2024_muni %>% filter(virus == "Dengue"),
  facet_var = "muni_name") +
  geom_vline(data = emergency_muni, aes(xintercept = epi_week), linetype = "dotted", color = "black", size = 1.5) +
  geom_vline(data = emergency_state_for_muni_plot, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1.5)

print(dengue_risk_muni)
ggsave(here("Graphs/Risk Levels of Dengue Outbreak in 2024 per Municipality with Emergency Dec.png"), dengue_risk_muni, height = 12, width = 16, dpi = 600, create.dir = TRUE)
```

### Mapping First Week Alert in 2024

```{r mapping-firstweekalert, fig.width=16, fig.height=12}
### Calculate First Alert Week
# This chunk calculates the first week each location entered risk level 2 or 3.

# States
alert_week_state <- risk_2024_state %>%
  filter(riskassessment %in% c("2", "3")) %>% # Filter for alert/epidemic levels
  group_by(location_name = uf, virus) %>%      # Group by state and virus
  summarise(
    first_week_alert_2024 = min(epi_week, na.rm = TRUE), # Find the earliest week
    .groups = 'drop'
  )

# Municipalities
alert_week_muni <- risk_2024_muni %>%
  filter(riskassessment %in% c("2", "3")) %>% # Filter for alert/epidemic levels
  group_by(location_name = muni_name, ibge, virus) %>% # Group by municipality, ibge, and virus
  summarise(
    first_week_alert_2024 = min(epi_week, na.rm = TRUE), # Find the earliest week
    .groups = 'drop'
  )

# Filter for virus data and rename columns for joining
dengue_alert_week_state_filtered <- alert_week_state %>%
 filter(virus == "Dengue") %>%
 rename(uf = location_name)

dengue_alert_week_muni_filtered <- alert_week_muni %>%
 filter(virus == "Dengue") %>%
  rename(municipality = location_name)

# Join the alert week data with the geospatial data
# States
dengue_map_alert_states <- states_sf %>%
 left_join(dengue_alert_week_state_filtered, by = c("abbrev_state" = "uf"))

# Capitals
dengue_map_alert_capitals <- capitals_points %>%
 left_join(dengue_alert_week_muni_filtered, by = c("code_muni" = "ibge"))

# Breaks
dengue_week_breaks <- c(5, 9, 12, Inf)

dengue_week_labels <- c("5-8 Weeks", "9-12 Weeks", "12+ Weeks")

dengue_map_alert_states <- dengue_map_alert_states %>%
 mutate(first_week_alert_2024 = as.numeric(first_week_alert_2024)) %>%
 mutate(alert_week_group = cut(first_week_alert_2024,
 breaks = dengue_week_breaks,
 labels = dengue_week_labels,
 include.lowest = TRUE,
 right = FALSE))
 
dengue_map_alert_capitals <- dengue_map_alert_capitals %>%
 mutate(first_week_alert_2024 = as.numeric(first_week_alert_2024)) %>%
 mutate(alert_week_group = cut(first_week_alert_2024,
 breaks = dengue_week_breaks,
 labels = dengue_week_labels,
 include.lowest = TRUE,
 right = FALSE))

# Generate the Map
#---------------------------------
dengue_risk_map_themed <- ggplot() +
 geom_sf(data = dengue_map_alert_states, aes(fill = alert_week_group), color = "black", linewidth = 0.5) +
 geom_sf(data = dengue_map_alert_capitals, aes(shape = alert_week_group),
 fill = NA, color = "darkgrey", alpha = 1, size = 4, stroke = 2) +
scale_fill_brewer(
    palette = "Oranges",
    direction = -1,
    name = "Risk Level >= 2 (States)",
 drop = F
 ) +
 scale_shape_manual(
 name = "Risk Level >= 2 (Capitals)",
         values = c(21, 22, 23),
  na.translate = FALSE,
 drop = F
 ) +
   guides(
    fill = guide_legend(order = 1),
    shape = guide_legend(order = 2)
  ) +
 theme_bw() +
 theme(
 legend.position = c(0.8, 0.1),
 legend.justification = c("left", "bottom"),
 legend.background = element_rect(fill = "transparent"),
 legend.box.background = element_blank(),
 panel.grid = element_blank(),
 panel.border = element_blank(),
 axis.title = element_text(),
 axis.text = element_text(),
 legend.title = element_text(size = 18), 
 legend.text = element_text(size = 18) 
 )

print(dengue_risk_map_themed)
ggsave(here("Graphs", "Map First Week Alert 2024.png"), dengue_risk_map_themed, width = 16, height = 12, dpi = 600)
```

### Mapping the lag between Alert and Emergency Declaration

```{r mapping-lag-diff, fig.width=16, fig.height=12}
# 1. Create the data frames for emergency declarations
emergency_week_state <- emergency_state %>%
  select(uf, declaration_week = epi_week)

# Select the IBGE code instead of the name for a reliable join
emergency_week_muni <- emergency_muni %>%
  select(ibge, declaration_week = epi_week)

# Define factor levels
lag_levels <- c(
  "Opportunistic",
  "Delayed",
  "No Declaration"
)

# States
lag_map_data_states <- states_sf %>%
  # Join alert data
  left_join(alert_week_state %>% filter(virus == "Dengue"), by = c("abbrev_state" = "location_name")) %>%
  # Join emergency declaration data (using the newly created df)
  left_join(emergency_week_state, by = c("abbrev_state" = "uf")) %>%
  mutate(
    # Calculate the difference in epidemiological weeks
    lag_weeks = declaration_week - first_week_alert_2024,
    
    # Create the categories based on the lag
    lag_category = case_when(
      is.na(declaration_week) ~ "No Declaration",
      lag_weeks == 0 ~ "Opportunistic",
      lag_weeks >=1 ~ "Delayed"
    ),
    
    # Convert to a factor to control the order in the plot legend
        lag_category = factor(lag_category, levels = lag_levels))

# Munis
lag_map_data_capitals <- capitals_points %>%
  # Join alert data (this join is already correct)
  left_join(alert_week_muni %>% filter(virus == "Dengue"), by = c("code_muni" = "ibge")) %>%
  # Join emergency declaration data BY IBGE CODE for reliability
  left_join(emergency_week_muni, by = c("code_muni" = "ibge")) %>%
  # Ensure we only have the selected capitals with data
  filter(!is.na(location_name)) %>%
  mutate(
    # Calculate the difference in epidemiological weeks
    lag_weeks = declaration_week - first_week_alert_2024,
    
    # Create the categories based on the lag
    lag_category = case_when(
      is.na(declaration_week) ~ "No Declaration",
      lag_weeks == 0 ~ "Opportunistic",
      lag_weeks >=1 ~ "Delayed"
    ),
    
    # Convert to a factor to control the order in the plot legend
       lag_category = factor(lag_category, levels = lag_levels))

# Generate the Map
lag_map <- ggplot() +
  geom_sf(data = lag_map_data_states, aes(fill = lag_category), color = "black", linewidth = 0.5) +
  geom_sf(data = lag_map_data_capitals, aes(shape = lag_category),
          fill = NA, color = "darkgrey", alpha = 1, size = 4, stroke = 2) +
  scale_fill_brewer(
    palette = "Paired",
    name = "Emergency Declaration (States)",
    drop = T, 
    na.translate = F
  ) +
  scale_shape_manual(
    name = "Emergency Declaration (Capitals)",
            values = c(21, 6, 23),
    drop = T, 
    na.translate = F
  ) +
   guides(
    fill = guide_legend(order = 1),
    shape = guide_legend(order = 2)
  ) +
  theme_bw() +
 theme(
  legend.position = c(0.8, 0.1),
  legend.justification = c("left", "bottom"),
  legend.background = element_rect(fill = "transparent"),
  legend.box.background = element_blank(),
  panel.grid = element_blank(),
  panel.border = element_blank(),
  axis.title = element_text(),
  axis.text = element_text(),
  legend.title = element_text(size = 18),
  legend.text = element_text(size = 18) 
 )

print(lag_map)
ggsave(here("Graphs", "Map_Lag_Alert_Declaration_2024.png"), lag_map, width = 16, height = 12, dpi = 600)
```

# Risk Duration and Emergency Declaration

## Prepare Data

```{r scatterplots-risk-duration, fig.width=16, fig.height=8}
# Prepare the data for state-level analysis
state_regression_data <- summary_table %>%
  filter(Level == "State") %>%
  mutate(declared_emergency = ifelse(Location %in% emergency_state$uf, 1, 0)) %>%
  select(
    location = Location,
    declared_emergency,
    max_consecutive_level2 = `Max. Consecutive Weeks (level 2)`,
    max_consecutive_level3 = `Max. Consecutive Weeks (level 3)`,
    max_consecutive_level2_or_3 = `Max. Consecutive Weeks (level 2 or 3)`
  )

# Create lookup tables for muni info and emergency declaration weeks
muni_info_lookup <- arboviruses_muni %>% distinct(muni_name, ibge, uf)
state_declaration_weeks <- emergency_state %>% select(uf, state_declaration_week = epi_week)
muni_declaration_weeks <- emergency_muni %>% select(ibge, muni_declaration_week = epi_week)

# Prepare the data for municipality-level regressions
muni_regression_data <- summary_table %>%
  filter(Level == "Municipality") %>%
  left_join(muni_info_lookup, by = c("Location" = "muni_name")) %>%
  left_join(state_declaration_weeks, by = "uf") %>%
  left_join(muni_declaration_weeks, by = "ibge") %>%
  mutate(
    declared_emergency = ifelse(ibge %in% emergency_muni$ibge, 1, 0),
    state_declared_before = ifelse(
      !is.na(state_declaration_week) & (is.na(muni_declaration_week) | state_declaration_week < muni_declaration_week),
      1, 0
    )
  ) %>%
  select(
    location = Location,
    declared_emergency,
    state_declared_before,
    max_consecutive_level2 = `Max. Consecutive Weeks (level 2)`,
    max_consecutive_level3 = `Max. Consecutive Weeks (level 3)`,
    max_consecutive_level2_or_3 = `Max. Consecutive Weeks (level 2 or 3)`
  )

# Add 'level' column to each and bind them
state_plot_data <- state_regression_data %>%
  mutate(level = "State")

muni_plot_data <- muni_regression_data %>%
  mutate(level = "Municipality")

scatter_plot_data <- bind_rows(state_plot_data, muni_plot_data) %>%
  mutate(
    # Convert 0/1 to a factor for plotting
    declared_emergency = as.factor(if_else(declared_emergency == 1, 
                                           "Declared Emergency", 
                                           "Did Not Declare"))
  )

# Separate data for the two plots
scatter_states <- scatter_plot_data %>% filter(level == "State")
scatter_munis <- scatter_plot_data %>% filter(level == "Municipality")
```

## Scatterplots

```{r scattering}
# Plot: States
plot_states <- ggplot(scatter_states, aes(x = max_consecutive_level2, y = max_consecutive_level3)) +
  geom_point(aes(color = declared_emergency), size = 3, alpha = 0.8) +
  # Add dotted lines for means
  geom_vline(aes(xintercept = mean(max_consecutive_level2, na.rm = TRUE)), linetype = "dotted", color = "black") +
  geom_hline(aes(yintercept = mean(max_consecutive_level3, na.rm = TRUE)), linetype = "dotted", color = "black") +
  # Add labels to identify states
  geom_text_repel(aes(label = location), size = 3.5, max.overlaps = Inf, 
                  color = "black", box.padding = 0.5) +
  scale_color_manual(values = c("Declared Emergency" = "royalblue", "Did Not Declare" = "grey"), name = NULL) +
  labs(
    x = "Max. Consecutive Weeks under Alert (Level 2)",
    y = "Max. Consecutive Weeks under Epidemic (Level 3)"
  ) +
  theme_bw(base_size = 14) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

# Plot: Munis
plot_munis <- ggplot(scatter_munis, aes(x = max_consecutive_level2, y = max_consecutive_level3)) +
  geom_point(aes(color = declared_emergency), size = 3, alpha = 0.8) +
  # Add dotted lines for means
  geom_vline(aes(xintercept = mean(max_consecutive_level2, na.rm = TRUE)), linetype = "dotted", color = "black") +
  geom_hline(aes(yintercept = mean(max_consecutive_level3, na.rm = TRUE)), linetype = "dotted", color = "black") +
  # Add labels to identify capitals
  geom_text_repel(aes(label = location), size = 3.5, max.overlaps = Inf,
                  color = "black", box.padding = 0.5) +
  scale_color_manual(values = c("Declared Emergency" = "royalblue", "Did Not Declare" = "grey"), name = NULL) +
  labs(
    x = "Max. Consecutive Weeks under Alert (Level 2)",
    y = "Max. Consecutive Weeks under Epidemic (Level 3)"
  ) +
  theme_bw(base_size = 14) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

# Print both plots
print(plot_states)
print(plot_munis)

# Save
ggsave(here("Graphs", "Scatter_Risk_Duration_States.png"),
  plot_states,
  width = 12, 
  height = 10,
  dpi = 600,
  create.dir = TRUE)

ggsave(here("Graphs", "Scatter_Risk_Duration_Munis.png"),
  plot_munis,
  width = 12,
  height = 10,
  dpi = 600,
  create.dir = TRUE)
```

# Logit Regressions: Risk Duration and Emergency Declaration

```{r running-logit}
# --- 1. State-Level Analysis ---
# Run the three logistic regression models for states
logit_state_l2 <- glm(declared_emergency ~ max_consecutive_level2, data = state_regression_data, family = binomial)
logit_state_l3 <- glm(declared_emergency ~ max_consecutive_level3, data = state_regression_data, family = binomial)
logit_state_l23 <- glm(declared_emergency ~ max_consecutive_level2_or_3, data = state_regression_data, family = binomial)


# --- 2. Municipality-Level Analysis ---
# Run the three logistic regression models for municipalities
logit_muni_l2 <- glm(declared_emergency ~ state_declared_before + max_consecutive_level2, data = muni_regression_data, family = binomial)
logit_muni_l3 <- glm(declared_emergency ~ state_declared_before + max_consecutive_level3, data = muni_regression_data, family = binomial)
logit_muni_l23 <- glm(declared_emergency ~ state_declared_before + max_consecutive_level2_or_3, data = muni_regression_data, family = binomial)

# --- 3. Create Table ---
# Combine the three state models into a single table
tbl_states_final <- tbl_merge(
  tbls = list(
    tbl_regression(logit_state_l2, exponentiate = TRUE),
    tbl_regression(logit_state_l3, exponentiate = TRUE),
    tbl_regression(logit_state_l23, exponentiate = TRUE)
  ),
  # Add column labels for each model
  tab_spanner = c("**Model 1**", "**Model 2**", "**Model 3**")
) %>%
  # Add a main title to the table
  modify_caption("Association Between Emergency Declaration and Risk Duration, States")

# Combine the three municipality models into a single table
tbl_munis_final <- tbl_merge(
  tbls = list(
    tbl_regression(logit_muni_l2, exponentiate = TRUE),
    tbl_regression(logit_muni_l3, exponentiate = TRUE),
    tbl_regression(logit_muni_l23, exponentiate = TRUE)
  ),
  tab_spanner = c("**Model 4**", "**Model 5**", "**Model 6**")
) %>%
  modify_caption("Association Between Emergency Declaration and Risk Duration, Municipalities")

# Display the final tables
tbl_states_final
tbl_munis_final

# Save
# Convert gtsummary tables to flextables to save them
ft_states <- as_flex_table(tbl_states_final)
ft_munis <- as_flex_table(tbl_munis_final)

# Create a Word document and add the tables
doc <- read_docx() %>%
  body_add_par("Logistic Regression Results for Dengue Emergency Declarations", style = "heading 1") %>%
  body_add_flextable(ft_states) %>%
  body_add_break() %>%
  body_add_flextable(ft_munis)

# Save the document
print(doc, target = "Regression_Tables.docx")
```

# Causal Impact

```{r causalimpact}
# Prepare Datasets 
country_data <- arboviruses_country %>% filter(virus == "Dengue", epi_year %in% c(2023, 2024)) %>% arrange(epi_year, epi_week)
state_data <- arboviruses_state %>% filter(virus == "Dengue", epi_year %in% c(2023, 2024)) %>% arrange(uf, epi_year, epi_week)
muni_data <- arboviruses_muni %>% filter(virus == "Dengue", epi_year %in% c(2023, 2024)) %>% arrange(muni_name, epi_year, epi_week)

# Function
run_causal_impact_analysis <- function(data_for_location, intervention_date_str, outcome_var) {
  
  # Create a complete time series with a real date index
  time_series_data <- data_for_location %>%
    mutate(date = as.Date(paste(substr(calendar_year_week, 1, 4), substr(calendar_year_week, 6, 7), 1, sep = "-"), "%Y-%U-%u"))
    
  # Define pre and post periods 
  intervention_date <- as.Date(paste(substr(intervention_date_str, 1, 4), substr(intervention_date_str, 6, 7), 1, sep = "-"), "%Y-%U-%u")
  pre_period  <- c(min(time_series_data$date), intervention_date)
  post_period <- c(intervention_date + 7, max(time_series_data$date))
  
  # Select only the outcome variable for the univariate model
  y <- time_series_data %>% select(!!sym(outcome_var))
  
  # Create the zoo object (just one column)
  impact_data_zoo <- zoo(y, time_series_data$date)
  
  # Run the model
  impact_result <- CausalImpact(impact_data_zoo, pre_period, post_period, model.args = list(niter = 5000, nseasons = 52))
  
  return(impact_result)
}
```

## Country
```{r causalimpactcountry}
# Define the national intervention date
fed_intervention_date <- emergency_country$calendar_year_week[1]

# --- Run for INCIDENCE ---
fed_impact_inc <- run_causal_impact_analysis(country_data, fed_intervention_date, "incidence")
p_inc <- plot(fed_impact_inc) +
  labs(title = "Emergency Declaration and Dengue Incidence - Brazil", subtitle = paste("Declaration on", fed_intervention_date), x = "Calendar Year-Week") +
  theme_bw(base_size = 16)
print(p_inc)
ggsave(here("Graphs", "CausalImpact_Federal_Brazil_Incidence.png"), p_inc, width = 15, height = 15, dpi = 600, create.dir = TRUE)

# Run for Death Rate
fed_impact_dth <- run_causal_impact_analysis(country_data, fed_intervention_date, "death_rate")
p_dth <- plot(fed_impact_dth) +
  labs(title = "Emergency Declaration and Dengue Death Rate - Brazil", subtitle = paste("Declaration on", fed_intervention_date), x = "Calendar Year-Week") +
  theme_bw(base_size = 16)
print(p_dth)
ggsave(here("Graphs", "CausalImpact_Federal_Brazil_DeathRate.png"), p_dth, width = 15, height = 15, dpi = 600, create.dir = TRUE)
```

## States

```{r causalimpactstates}
# Get a list of states that declared an emergency
states_to_analyze <- unique(emergency_state$uf)

# Loop through each state to run the analysis
for (state_code in states_to_analyze) {

  # Get the intervention date for this state
  state_intervention_date <- emergency_state$calendar_year_week[emergency_state$uf == state_code][1]
  
  # Filter the data for this state
  data_for_state <- state_data %>% filter(uf == state_code)

  # --- Run for INCIDENCE ---
  impact_inc <- run_causal_impact_analysis(data_for_state, state_intervention_date, "incidence")
  p_inc <- plot(impact_inc) +
    labs(title = paste("Emergency Declaration and Dengue Incidence -", state_code), subtitle = paste("Declaration on", state_intervention_date), x = "Calendar Year-Week") +
    theme_bw(base_size = 16)
  print(p_inc)
  ggsave(here("Graphs", paste0("CausalImpact_State_", state_code, "_Incidence.png")), p_inc, width = 15, height = 15, dpi = 600, create.dir = TRUE)
  
  # --- Run for DEATH RATE ---
  impact_dth <- run_causal_impact_analysis(data_for_state, state_intervention_date, "death_rate")
  p_dth <- plot(impact_dth) +
    labs(title = paste("Emergency Declaration and Dengue Death Rate -", state_code), subtitle = paste("Declaration on", state_intervention_date), x = "Calendar Year-Week") +
    theme_bw(base_size = 16)
  print(p_dth)
  ggsave(here("Graphs", paste0("CausalImpact_State_", state_code, "_DeathRate.png")), p_dth, width = 15, height = 15, dpi = 600, create.dir = TRUE)
}
```

## Municipalities

```{r causalimpactmunis}
# Get a list of municipalities that declared an emergency
munis_to_analyze <- emergency_muni %>% 
filter(muni_name == "FLORIANOPOLIS") %>%
select(ibge) %>%
  pull() %>%
  unique()

# Loop through each municipality to run the analysis
for (muni_code in munis_to_analyze) {
  
  # Get the municipality name and intervention date
  muni_info <- emergency_muni %>% filter(ibge == muni_code)
  muni_name_str <- muni_info$muni_name[1]
  muni_intervention_date <- muni_info$calendar_year_week[1]
  
  # Filter the data for this municipality
  data_for_muni <- muni_data %>% filter(muni_name == muni_name_str)
  
  # --- Run for INCIDENCE ---
  impact_inc <- run_causal_impact_analysis(data_for_muni, muni_intervention_date, "incidence")
  p_inc <- plot(impact_inc) +
    labs(title = paste("Emergency Declaration and Dengue Incidence -", muni_name_str), subtitle = paste("Declaration on", muni_intervention_date), x = "Calendar Year-Week") +
    theme_bw(base_size = 16)
  print(p_inc)
  ggsave(here("Graphs", paste0("CausalImpact_Muni_", muni_name_str, "_Incidence.png")), p_inc, width = 15, height = 15, dpi = 600, create.dir = TRUE)
  
  # --- Run for DEATH RATE ---
  impact_dth <- run_causal_impact_analysis(data_for_muni, muni_intervention_date, "death_rate")
  p_dth <- plot(impact_dth) +
    labs(title = paste("Emergency Declaration and Dengue Death Rate -", muni_name_str), subtitle = paste("Declaration on", muni_intervention_date), x = "Calendar Year-Week") +
    theme_bw(base_size = 16)
  print(p_dth)
  ggsave(here("Graphs", paste0("CausalImpact_Muni_", muni_name_str, "_DeathRate.png")), p_dth, width = 15, height = 15, dpi = 600, create.dir = TRUE)
}
```



