---
title: "Replication: Gaps in Early Warning Systems (EWSs) during  Brazil`s Largest Dengue Epidemic"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
lang: pt-BR
output:
  html_document:
    number_sections: true
    toc: true
    code_folding: hide
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
if (.Platform$OS.type == "windows") {
  Sys.setlocale("LC_CTYPE", "Portuguese_Brazil.1252")
} else {
  Sys.setlocale("LC_CTYPE", "pt_BR.UTF-8")
}

library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(ggsci)
library(here)
library(haven)
library(readr)
library(stringr)
library(officer)
library(flextable)
library(geobr)
library(sf)
library(scales)
library(RColorBrewer)
```

---

# Data Preparation

## Gathering Data

```{r gather-data}
# All arbovirsuses data:
arboviruses_full <- read_csv(here("Data/cases_all_virus_2014_2024.csv")) %>% 
  rename(calendar_year = year,
         calendar_week = week)  %>%
  mutate(virus = str_to_title(virus),
          calendar_year_week = paste0(calendar_year, "w", calendar_week)) 

# Epidemiological week starting on the first week of the winter
epiweek40 <- read_dta(
  here("Data/epidemiological_week_startsweek40.dta")) %>% 
  rename(calendar_year_week = wy,
         epi_week = epidemiologic_week_startsweek40,
         epi_year = year_week40) %>% 
  mutate(epi_week = as.numeric(epi_week),
         epi_year = as.numeric(epi_year)) %>% 
  mutate(epi_year_week = paste0(epi_year, "w", epi_week))

#List of municipalities of interest
codibge <- list(
  "Rio Branco"      = "120040",
  "Maceió"          = "270430",
  "Macapá"          = "160030",
  "Manaus"          = "130260",
  "Salvador"        = "292740",
  "Fortaleza"       = "230440",
  "Brasília"        = "530010",
  "Vitória"         = "320530",
  "Goiânia"         = "520870",
  "São Luís"        = "211130",
  "Cuiabá"          = "510340",
  "Campo Grande"    = "500270",
  "Belo Horizonte"  = "310620",
  "Belém"           = "150140",
  "João Pessoa"     = "250750",
  "Curitiba"        = "410690",
  "Recife"          = "261160",
  "Teresina"        = "221100",
  "Rio de Janeiro"  = "330455",
  "Natal"           = "240810",
  "Porto Alegre"    = "431490",
  "Porto Velho"     = "110020",
  "Boa Vista"       = "140010",
  "Florianópolis"   = "420540",
  "São Paulo"       = "355030",
  "Aracaju"         = "280030",
  "Palmas"          = "172100")
```

---

## Preparing Data

```{r preparing-data}
# 1. Create the base dataset
arboviruses_base <- arboviruses_full %>%
  select(ibge, calendar_year_week, municipality, uf, virus, likely_cases, deaths, population) %>%
  full_join(epiweek40, by = "calendar_year_week") %>%
  arrange(epi_year, epi_week) %>%
  filter(!is.na(municipality), !is.na(epi_year), !is.na(epi_week))

# 2. Calculate population for each epidemiological year
# States
state_population <- arboviruses_base %>%
  distinct(uf, ibge, epi_year, .keep_all = TRUE) %>%
  group_by(uf, epi_year) %>%
  summarise(state_pop = sum(population), .groups = "drop")

# Country
country_population <- arboviruses_base %>%
  distinct(ibge, epi_year, .keep_all = TRUE) %>%
  group_by(epi_year) %>%
  summarise(country_pop = sum(population), .groups = "drop")

# 3. Create datasets for each geographic level

# Munis (Individual Viruses)
arboviruses_muni_ind <- arboviruses_base %>%
  filter(ibge %in% unlist(codibge)) %>%
  group_by(ibge, municipality, uf, epi_year, epi_week, epi_year_week, virus) %>%
  summarise(
    cases = sum(likely_cases),
    deaths = sum(deaths),
    population = first(population), # pop is measured on year basis
    .groups = "drop"
  ) %>%
  mutate(
    incidence = (cases / population) * 100000,
    death_rate = (deaths / population) * 100000
  ) %>%
  # Calculate cumulative per epi year
  group_by(ibge, municipality, uf, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

# States (Individual Viruses)
arboviruses_state_ind <- arboviruses_base %>%
  group_by(uf, epi_year, epi_week, epi_year_week, virus) %>%
  summarise(
    cases = sum(likely_cases),
    deaths = sum(deaths),
    .groups = "drop"
  ) %>%
  full_join(state_population, by = c("uf", "epi_year")) %>%
  mutate(
    incidence = (cases / state_pop) * 100000,
    death_rate = (deaths / state_pop) * 100000
  ) %>%
  # Calculate cumulative per epi year
  group_by(uf, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

# Country (Individual Viruses)
arboviruses_country_ind <- arboviruses_base %>%
  group_by(epi_year, epi_week, epi_year_week, virus) %>%
  summarise(
    cases = sum(likely_cases),
    deaths = sum(deaths),
    .groups = "drop"
  ) %>%
  left_join(country_population, by = "epi_year") %>%
  mutate(
    incidence = (cases / country_pop) * 100000,
    death_rate = (deaths / country_pop) * 100000,
    country = "Brazil"
  ) %>%
  # Calculate cumulative per epi year
  group_by(country, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

# 4. Create data for the sum of arboviruses

# Muni (Sum of Viruses)
all_virus_muni <- arboviruses_muni_ind %>%
  group_by(ibge, municipality, uf, epi_year, epi_week, epi_year_week) %>%
  summarise(
    cases = sum(cases),
    deaths = sum(deaths),
    population = first(population),
    .groups = "drop"
  ) %>%
  mutate(
    incidence = (cases / population) * 100000,
    death_rate = (deaths / population) * 100000,
    virus = "Sum of Arboviruses"
  ) %>%
  # Calculate cumulative per epi year
  group_by(ibge, municipality, uf, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

# States (Sum of Viruses)
all_virus_state <- arboviruses_state_ind %>%
  group_by(uf, epi_year, epi_week, epi_year_week) %>%
  summarise(
    cases = sum(cases),
    deaths = sum(deaths),
    .groups = "drop"
  ) %>%
  left_join(state_population, by = c("uf", "epi_year")) %>%
  mutate(
    incidence = (cases / state_pop) * 100000,
    death_rate = (deaths / state_pop) * 100000,
    virus = "Sum of Arboviruses"
  ) %>%
  # Calculate cumulative per epi year
  group_by(uf, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

# Country (Sum of Viruses)
all_virus_country <- arboviruses_country_ind %>%
  group_by(epi_year, epi_week, epi_year_week) %>%
  summarise(
    cases = sum(cases),
    deaths = sum(deaths),
    .groups = "drop"
  ) %>%
  left_join(country_population, by = "epi_year") %>%
  mutate(
    incidence = (cases / country_pop) * 100000,
    death_rate = (deaths / country_pop) * 100000,
    country = "Brazil",
    virus = "Sum of Arboviruses"
  ) %>%
  # Calculate cumulative per epi year
  group_by(country, virus, epi_year) %>%
  arrange(epi_week, .by_group = TRUE) %>%
  mutate(
    cumulative_incidence = cumsum(incidence),
    cumulative_death_rate = cumsum(death_rate)
  ) %>%
  ungroup()

# 5. Bind the datasets
arboviruses_muni <- bind_rows(arboviruses_muni_ind, all_virus_muni) %>%
  mutate(muni_name = stringr::str_remove(municipality, "^\\d+\\s"))
arboviruses_state <- bind_rows(arboviruses_state_ind, all_virus_state)
arboviruses_country <- bind_rows(arboviruses_country_ind, all_virus_country)
```

---

# Data Visualization

---

# Plotting Function

```{r funct-plot}
create_arbovirus_plot <- function(data, y_var, virus_filter_expr, y_lab,
                                  grouping_var, legend_position,
                                  facet_by = NULL,
                                  year_interval = 1,
                                  file_name) {

  # Use quosures for tidy evaluation
  y_var_quo <- enquo(y_var)
  grouping_var_quo <- enquo(grouping_var)
  virus_filter_quo <- enquo(virus_filter_expr)

  # Find the first week of each epidemiological year to use as a label
  axis_breaks <- data %>%
    group_by(epi_year) %>%
    slice_head(n = 1) %>%
        ungroup() %>%
    slice(seq(1, n(), by = year_interval)) %>%
    pull(epi_year_week)

  # Build the plot
  p <- data %>%
    arrange(epi_year, epi_week) %>%
    filter(!!virus_filter_quo) %>%
    ggplot(aes(x = epi_year_week, y = !!y_var_quo, group = !!grouping_var_quo, color = virus)) +
    geom_line(alpha = 0.8, linewidth = 0.9) +
    scale_color_lancet(name = "Virus") +
    labs(
      x = "Epidemiological Year-Week",
      y = y_lab
    ) +
    scale_x_discrete(breaks = axis_breaks) +
    theme_classic(base_size = 14) + 
    theme(
      axis.title = element_text(face = "bold", size = rel(1)), 
      legend.position = legend_position,
      legend.title = element_text(face = "bold", size = rel(1)),
      legend.text = element_text(size = rel(1)),
      panel.grid.major.y = element_line(color = "grey85", linewidth = 0.5),
      axis.text.x = element_text(angle = 45, hjust = 1)
    )

  # Add faceting if specified
  if (!is.null(facet_by)) {
    p <- p + facet_wrap(vars(!!sym(facet_by)), scales = "free_y")
  }

  # Print and save
  print(p)
  ggsave(
    here("Graphs", file_name),
    p,
    height = 15,
    width = 15,
    dpi = 300,
    create.dir = TRUE
  )
}
```

---

## Incidence

### Country-wide Incidence

#### Dengue
```{r incidence-country-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_country,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = epi_year,
  legend_position = "top",
  file_name = "Incidence of Dengue in Brazil.png"
)
```

#### Dengue, Chikungunya, and Zika
```{r incidence-country-all, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_country,
  y_var = incidence,
  virus_filter_expr = virus != "Sum of Arboviruses",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "top",
  file_name = "Incidence of All Arboviruses in Brazil.png"
)
```

### State-level Incidence

#### Dengue
```{r incidence-state-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_state,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = 1,
  legend_position = "top",
  facet_by = "uf",
  year_interval = 2,
  file_name = "Incidence of Dengue per State.png"
)
```

#### Dengue, Chikungunya, and Zika
```{r incidence-state-all, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_state,
  y_var = incidence,
  virus_filter_expr = virus == "Sum of Arboviruses",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "top",
  facet_by = "uf",
  year_interval = 2,
  file_name = "Incidence of All Arboviruses per State.png"
)
```

### Selected Municipalities Incidence

#### Dengue
```{r incidence-muni-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_muni,
  y_var = incidence,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "top",
  facet_by = "muni_name",
  year_interval = 2,
  file_name = "Incidence of Dengue per Selected Municipalities.png"
)
```

#### Dengue, Chikungunya, and Zika
```{r incidence-muni-all, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_muni,
  y_var = incidence,
  virus_filter_expr = virus == "Sum of Arboviruses",
  y_lab = "Incidence Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "top",
  facet_by = "muni_name",
  year_interval = 2,
  file_name = "Incidence of All Arboviruses per Selected Municipalities.png"
)
```

---

## Mortality

### Country-wide Deaths

#### Dengue
```{r deaths-country-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_country,
  y_var = death_rate,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Death Rate (per 100,000 pop.)",
  grouping_var = epi_year,
  legend_position = "top",
  file_name = "Deaths of Dengue in Brazil.png"
)
```

#### Dengue, Chikungunya, and Zika
```{r deaths-country-all, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_country,
  y_var = death_rate,
  virus_filter_expr = virus != "Sum of Arboviruses",
  y_lab = "Death Rate (per 100,000 pop.)",
  grouping_var = virus,
  legend_position = "top",
  file_name = "Deaths of All Arboviruses in Brazil.png"
)
```

### State-level Deaths

#### Dengue
```{r deaths-state-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_state,
  y_var = cumulative_death_rate,
  virus_filter_expr = virus == "Dengue",
  y_lab = "Cumulative Death Rate (per 100,000 pop.) per Epidemiological Year",
  grouping_var = 1,
  legend_position = "top",
  facet_by = "uf",
  year_interval = 2,
  file_name = "Deaths of Dengue per State.png"
)
```

#### Dengue, Chikungunya, and Zika
```{r deaths-state-all, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_state,
  y_var = cumulative_death_rate,
  virus_filter_expr = virus == "Sum of Arboviruses",
  y_lab = "Cumulative Death Rate (per 100,000 pop.) per Epidemiological Year",
  grouping_var = virus,
  legend_position = "top",
  facet_by = "uf",
  year_interval = 2,
  file_name = "Deaths of All Arboviruses per State.png"
)
```

### Selected Municipalities Deaths

#### Dengue
```{r deaths-muni-dengue, fig.width=15, fig.height=15}
create_arbovirus_plot(
  data = arboviruses_muni,
  y_var = cumulative_death_rate,
 virus_filter_expr = virus == "Dengue",
  y_lab = "Cumulative Death Rate (per 100,000 pop.) per Epidemiological Year",
  grouping_var = 1,
  legend_position = "top",
  facet_by = "muni_name",
  year_interval = 2,
  file_name = "Deaths of Dengue per Selected Municipalities.png"
)
```

#### Dengue, Chikungunya, and Zika
```{r deaths-muni-all, fig.width=15, fig.height=10}
create_arbovirus_plot(
  data = arboviruses_muni,
  y_var = cumulative_death_rate,
  virus_filter_expr = virus == "Sum of Arboviruses",
  y_lab = "Cumulative Death Rate (per 100,000 pop.) per Epidemiological Year",
  grouping_var = virus,
  legend_position = "top",
  facet_by = "muni_name",
  year_interval = 2,
  file_name = "Deaths of All Arboviruses per Selected Municipalities.png"
)
```

---

## Mapping Incidence and Mortality in 2024

```{r spatial}
# Get Geospatial Data
#---------------------------------
states_sf <- read_state(year = 2020)
capitals_points <- read_municipality(year = 2020) %>%
  st_centroid() %>%
  mutate(code_muni = as.double(substr(as.character(code_muni), 1, 6)))
```

### Dengue

```{r mapping-dengue, fig.width=12, fig.height=10}
# Define the Epidemiological Year of Interest
epi_year_to_plot <- 2024

# Combine Data with Map
dengue_states_map_data <- arboviruses_state %>%
  filter(epi_year == epi_year_to_plot,
         virus == "Dengue") %>%
  slice_max(order_by = epi_week, n = 1) %>%
  left_join(states_sf, ., by = c("abbrev_state" = "uf"))

dengue_capitals_map_data <- arboviruses_muni %>%
  filter(epi_year == epi_year_to_plot,
         virus == "Dengue",
         ibge %in% codibge) %>%
  slice_max(order_by = epi_week, n = 1) %>%
  left_join(capitals_points, ., by = c("code_muni" = "ibge")) %>% 
  filter(!is.na(municipality), !is.na(epi_year), !is.na(epi_week))

# Helper function to create clean labels for plot
create_clean_labels <- function(breaks) {
  labels <- c()
  for (i in 1:(length(breaks) - 1)) {
    start <- formatC(breaks[i], format = "f", digits = 2, big.mark = ",")
    end <- formatC(breaks[i+1], format = "f", digits = 2, big.mark = ",")
    labels[i] <- paste0(start, " - ", end)
  }
  return(labels)
}

# Define breaks for Incidence
inc_breaks <- c(0, 500, 2500, Inf) 
inc_labels <- c("0 - 499", "500 - 2,499", "2,500+")

# Define breaks for Death Rate
dengue_death_rate_breaks <- c(0, 2, 5, Inf) 
dengue_death_rate_labels <- c("0 - 1.9", "2 - 4.9", "5+")

# Apply these unified breaks and clean labels to both datasets
dengue_states_map_data <- dengue_states_map_data %>%
  mutate(
    incidence_cat = cut(cumulative_incidence, breaks = inc_breaks, labels = inc_labels, include.lowest = TRUE, right = FALSE),
    death_rate_cat = cut(cumulative_death_rate, breaks = dengue_death_rate_breaks, labels = dengue_death_rate_labels, include.lowest = TRUE, right = FALSE)
  )
dengue_capitals_map_data <- dengue_capitals_map_data %>%
  mutate(
    incidence_cat = cut(cumulative_incidence, breaks = inc_breaks, labels = inc_labels, include.lowest = TRUE, right = FALSE),
    death_rate_cat = cut(cumulative_death_rate, breaks = dengue_death_rate_breaks, labels = dengue_death_rate_labels, include.lowest = TRUE, right = FALSE)
  )

# Generate the Maps
#---------------------------------

### MAP 1: CUMULATIVE INCIDENCE

dengue_incidence_map_themed <- ggplot() +
  geom_sf(data = dengue_states_map_data, aes(fill = incidence_cat), color = "black", linewidth = 0.5) +
  geom_sf(data = dengue_capitals_map_data, aes(shape = incidence_cat), 
         fill = NA, color = "darkgrey", alpha = 0.7, size = 4, stroke = 2) +
   scale_fill_brewer(
    palette = "Blues",
    name = "Dengue Incidence (States)\n(per 100,000 pop.)",
    na.translate = FALSE,
    drop = TRUE 
  ) +
  scale_shape_discrete(
    name = "Dengue Incidence (Capitals)\n(per 100,000 pop.)",
    drop = FALSE
  ) +
  theme_bw() +
  theme(
    legend.position = c(0.8, 0.1),
    legend.justification = c("left", "bottom"),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.title = element_text(),
    axis.text = element_text()
  )

print(dengue_incidence_map_themed)
ggsave(here("Graphs", paste0("Map Dengue Incidence ", epi_year_to_plot, ".png")), dengue_incidence_map_themed, width = 12, height = 10, dpi = 300)

### MAP 2: CUMULATIVE DEATH RATE

dengue_deaths_rate_map_themed <- ggplot() +
  geom_sf(data = dengue_states_map_data, aes(fill = death_rate_cat), color = "black", linewidth = 0.5) +
  geom_sf(data = dengue_capitals_map_data, aes(shape = death_rate_cat), 
           fill = NA, color = "darkgrey", alpha = 0.7, size = 4, stroke = 2) +
  scale_fill_brewer(
    palette = "Reds",
    name = "Dengue Death Rate (States)\n(per 100,000 pop.)",
    na.translate = FALSE,
    drop = TRUE
  ) +
  scale_shape_discrete(
    name = "Dengue Death Rate (Capitals)\n(per 100,000 pop.)",
    drop = FALSE
  ) +
  theme_bw() +
  theme(
    legend.position = c(0.8, 0.1),
    legend.justification = c("left", "bottom"),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    axis.title = element_text(),
    axis.text = element_text()
  )

print(dengue_deaths_rate_map_themed)
ggsave(here("Graphs", paste0("Map Dengue Death ", epi_year_to_plot, ".png")), dengue_deaths_rate_map_themed, width = 12, height = 10, dpi = 300)
```

# Risk Assessment

## Risk assessment function

```{r risk-function}
# Helper function (unchanged)
get_location_name <- function(df) {
  if ("muni_name" %in% names(df)) {
    return(paste("municipality:", paste(unique(df$muni_name), collapse = ", ")))
  } else if ("uf" %in% names(df)) {
    return(paste("state:", paste(unique(df$uf), collapse = ", ")))
  } else if ("country" %in% names(df)) {
    return(paste("country:", paste(unique(df$country), collapse = ", ")))
  } else {
    return("unknown location")
  }
}

risk_analysis <- function(df, analysis_year = 2024) {
  
  # === PART 1: PREPARE DATA & BUILD ENDEMIC CHANNEL ===
  
  # Define the historical period for the endemic channel (last 5 years)
  historic_years <- (analysis_year - 5):(analysis_year - 1)
  
  # Create the endemic channel based on the last 5 years 
  endemic_channel <- df %>%
    filter(epi_year %in% historic_years) %>%
    group_by(epi_week) %>%
    summarise(
      median_incidence = median(incidence),
      high = quantile(incidence, 0.75),
      low = quantile(incidence, 0.25),
      .groups = "drop"
    )
  
  # Isolate the current and previous year's data for comparison
  current_year_data <- df %>%
    filter(epi_year == analysis_year) %>%
    select(epi_week, incidence_current = incidence, deaths_current = deaths)
  
  previous_year_data <- df %>%
    filter(epi_year == (analysis_year - 1)) %>%
    select(epi_week, incidence_previous = incidence)
  
  # Combine all the data into a single data frame for assessment
  risk_final <- tibble(epi_week = sort(unique(df$epi_week))) %>%
    left_join(endemic_channel, by = "epi_week") %>%
    left_join(current_year_data, by = "epi_week") %>%
    left_join(previous_year_data, by = "epi_week") %>%
    # Replace NA values with 0 to allow for calculations
    mutate(
      across(c(incidence_current, deaths_current, incidence_previous), ~replace_na(., 0)),
      across(c(median_incidence, high, low), ~replace_na(., 0)) # Handle NAs in the channel
    )
  
  # === PART 2: APPLY RISK ASSESSMENT RULES ===
  
  risk_final <- risk_final %>%
    arrange(epi_week) %>%
    mutate(
      # Condition 1: Current incidence is within the endemic channel
      incidence_within_channel = incidence_current >= low & incidence_current <= high,
      
      # Condition 2: A sustained increase in cases compared to the previous year
      increased_vs_last_year = ifelse(incidence_current > incidence_previous, 1, 0),
      incidence_increase_4w_vs_last_year = zoo::rollapply(
        increased_vs_last_year, 4,
        function(x) as.integer(all(x == 1)),
        fill = 0, align = "right"),
      
      # Condition 3: Current incidence is above the high threshold of the endemic channel
      incidence_above_channel = incidence_current > high,
      
      # Condition 4: Deaths have been confirmed in the current week
      deaths_confirmed = deaths_current > 0
    ) %>%
    # Assign a risk level based on a nested, hierarchical logic
    mutate(
      riskassessment = case_when(
      
      # EMERGENCY - Confirmed deaths AND incidence above the endemic channel
      deaths_confirmed & incidence_above_channel ~ 3L,
      
      # ALERT - Confirmed deaths AND EITHER incidence is within the endemic channel but no deaths OR increase in incidence for four consecutive weeks compared to the previous year
      deaths_confirmed & (incidence_within_channel | incidence_increase_4w_vs_last_year == 1) ~ 2L,
      
      # INITIAL RESPONSE - No deaths AND EITHER incidence is within the endemic channel OR increase in incidence for four consecutive weeks compared to the previous year
!deaths_confirmed & (incidence_within_channel | incidence_increase_4w_vs_last_year == 1) ~ 1L,
   
        # Default: LOW RISK
        TRUE ~ 0L
      )
    )
  
    return(list(risk_final = risk_final))
}
```

### Running Risk assessment function

```{r risk-func-run}
# Country Level
nested_data_country <- arboviruses_country %>%
  group_by(country, virus) %>%
  nest()

results_country <- nested_data_country %>%
  mutate(risk_analysis_results = map(data, risk_analysis))

results_country_filtered <- results_country %>%
  filter(!map_lgl(risk_analysis_results, is.null))

risk_2024_country <- results_country_filtered %>%
  mutate(risk_data = map(risk_analysis_results, ~ .x$risk_final)) %>%
  select(country, virus, risk_data) %>%
  unnest(cols = c(risk_data))

# State Level
nested_data_state <- arboviruses_state %>%
  group_by(uf, virus) %>%
  nest()

results_state <- nested_data_state %>%
  mutate(risk_analysis_results = map(data, risk_analysis))

results_state_filtered <- results_state %>%
  filter(!map_lgl(risk_analysis_results, is.null))
  
risk_2024_state <- results_state_filtered %>%
  mutate(risk_data = map(risk_analysis_results, ~ .x$risk_final)) %>%
  select(uf, virus, risk_data) %>%
  unnest(cols = c(risk_data))

# Municipality Level
nested_data_muni <- arboviruses_muni %>%
  group_by(municipality, muni_name, virus, ibge) %>%
  nest()

results_muni <- nested_data_muni %>%
  mutate(risk_analysis_results = map(data, risk_analysis))

results_muni_filtered <- results_muni %>%
  filter(!map_lgl(risk_analysis_results, is.null))

risk_2024_muni <- results_muni_filtered %>%
  mutate(risk_data = map(risk_analysis_results, ~ .x$risk_final)) %>%
  select(municipality, muni_name, ibge, virus, risk_data) %>%
  unnest(cols = c(risk_data))
```

### Summary Table of Risk Assessment

```{r table-risk-summary}
# --- Country Level Table ---

# Calculate first alert week
alert_week_country <- risk_2024_country %>%
  filter(virus %in% c("Dengue", "Sum of Arboviruses")) %>%
  group_by(location_name = country, virus) %>%
  summarise(first_week_alert_2024 = min(epi_week[riskassessment >= 2], na.rm = TRUE), .groups = 'drop') %>%
  mutate(first_week_alert_2024 = as.numeric(first_week_alert_2024)) %>%
  mutate(first_week_alert_2024 = na_if(first_week_alert_2024, Inf))

# Display the table
knitr::kable(alert_week_country,
             caption = "Country Level: First Alert Week in 2024 (Level >= 2)",
             col.names = c("Country", "Virus", "2024 First Alert Week"))


# --- State Level Table ---

# Calculate first alert week
alert_week_state <- risk_2024_state %>%
  filter(virus %in% c("Dengue", "Sum of Arboviruses")) %>%
  group_by(location_name = uf, virus) %>%
  summarise(first_week_alert_2024 = min(epi_week[riskassessment >= 2], na.rm = TRUE), .groups = 'drop') %>%
  mutate(first_week_alert_2024 = as.numeric(first_week_alert_2024)) %>%
  mutate(first_week_alert_2024 = na_if(first_week_alert_2024, Inf))

# Display the table
knitr::kable(alert_week_state,
             caption = "State Level: First Alert Week in 2024 (Level >= 2)",
             col.names = c("UF", "Virus", "2024 First Alert Week"))


# --- Municipality Level Table ---

# Calculate first alert week
alert_week_muni <- risk_2024_muni %>%
  filter(virus %in% c("Dengue", "Sum of Arboviruses")) %>%
  group_by(location_name = muni_name, ibge, virus) %>%
  summarise(first_week_alert_2024 = min(epi_week[riskassessment >= 2], na.rm = TRUE), .groups = 'drop') %>%
  mutate(first_week_alert_2024 = as.numeric(first_week_alert_2024)) %>%
  mutate(first_week_alert_2024 = na_if(first_week_alert_2024, Inf))

# Display the table
knitr::kable(alert_week_muni,
             caption = "Selected Municipalities: First Alert Week in 2024 (Level >= 2)",
             col.names = c("Municipality", "IBGE", "Virus", "2024 First Alert Week"))


# Create word docx
doc <- read_docx()

# --- Country Level ---
doc <- doc %>%
  body_add_par("Country Level: First Alert Week in 2024 (Level >= 2)", style = "heading 1") %>%
  body_add_flextable(flextable(alert_week_country)) %>%
  body_add_par("")

# --- State Level ---
doc <- doc %>%
  body_add_par("State Level: First Alert Week in 2024 (Level >= 2)", style = "heading 1") %>%
  body_add_flextable(flextable(alert_week_state)) %>%
  body_add_par("")

# --- Municipality Level ---
doc <- doc %>%
  body_add_par("Municipality Level: First Alert Week in 2024 (Level >= 2)", style = "heading 1") %>%
  body_add_flextable(flextable(alert_week_muni)) %>%
  body_add_par("")

# Save
print(doc, target = "epidemic_tables.docx")
```

### Combining with Emergency Declarations in Epidemiological Year 2024

```{r preparing-risk-data}
# Country Risk Assessment
risk_2024_country <- risk_2024_country %>%
  mutate(epi_week = as.numeric(epi_week),
         riskassessment = as.factor(riskassessment))

# States Risk Assessment
risk_2024_state <- risk_2024_state %>%
  mutate(epi_week = as.numeric(epi_week),
         riskassessment = as.factor(riskassessment))

# Municipalities Risk Assessment
risk_2024_muni <- risk_2024_muni %>%
  mutate(epi_week = as.numeric(epi_week),
         riskassessment = as.factor(riskassessment))

# Creation of the Centro de Operações de Emergências de Saúde Pública para Dengue e outras Arboviroses, no âmbito do Ministério da Saúde
emergency_country <- tibble::tibble(
  label = "02/02/2024",
  date_issued = dmy("02/02/2024")
) %>%
  mutate(
    year = year(date_issued),
    week = week(date_issued), # Get standard week number
    calendar_year_week = paste0(year, "w", as.integer(week)) # Create a join key (e.g., "2024w5")
  ) %>%
  left_join(epiweek40, by = "calendar_year_week") 

# State Emergency Declarations
emergency_state <- tibble::tribble(
    ~uf, ~date_issued,
    "AC", "05/01/2024",
    "DF", "25/01/2024",
    "GO", "02/02/2024",
    "ES", "21/02/2024",
    "RJ", "21/02/2024",
    "SC", "22/02/2024", # Same day as Florianópolis
    "MG", "26/02/2024",
    "AP", "28/02/2024",
    "SP", "05/03/2024",
    "RS", "12/03/2024",
    "PR", "14/03/2024"
  ) %>%
  mutate(
    date_issued = dmy(date_issued),
    year = year(date_issued),
    week = week(date_issued), # Get standard week number
    calendar_year_week = paste0(year, "w", as.integer(week)) # Create a join key
  ) %>%
  left_join(epiweek40, by = "calendar_year_week") 

# Municipal Emergency Declarations
emergency_muni <- tibble::tribble(
    ~muni_name, ~date_issued,
    "FLORIANOPOLIS", "29/11/2023", 
    "RIO DE JANEIRO", "02/02/2024",
    "BELO HORIZONTE", "16/02/2024",
    "FLORIANOPOLIS", "22/02/2024", # Same day as SC
    "NATAL", "01/03/2024",
    "GOIANIA", "13/03/2024",
    "SAO PAULO", "18/03/2024",
    "PORTO ALEGRE", "22/04/2024"
  ) %>%
  mutate(
    date_issued = dmy(date_issued),
    year = year(date_issued),
    week = week(date_issued), # Get standard week number
    calendar_year_week = paste0(year, "w", as.integer(week)) # Create a join key
  ) %>%
  left_join(epiweek40, by = "calendar_year_week")

# Aggregate labels for states declaring emergency in the same week
emergency_state_labels <- emergency_state %>%
  group_by(epi_week) %>%
  summarise(uf_labels = paste(uf, collapse = "\n"))

# Create the state emergency data needed for the municipal plots
muni_to_state_map <- tibble::tribble(
  ~muni_name, ~uf,
  "RIO BRANCO", "AC", "MACEIO", "AL", "MACAPA", "AP", "MANAUS", "AM", "SALVADOR", "BA",
  "FORTALEZA", "CE", "BRASILIA", "DF", "VITORIA", "ES", "GOIANIA", "GO", "SAO LUIS", "MA",
  "CUIABA", "MT", "CAMPO GRANDE", "MS", "BELO HORIZONTE", "MG", "BELEM", "PA",
  "JOAO PESSOA", "PB", "CURITIBA", "PR", "RECIFE", "PE", "TERESINA", "PI",
  "RIO DE JANEIRO", "RJ", "NATAL", "RN", "PORTO ALEGRE", "RS", "PORTO VELHO", "RO",
  "BOA VISTA", "RR", "FLORIANOPOLIS", "SC", "SAO PAULO", "SP", "ARACAJU", "SE", "PALMAS", "TO"
)

emergency_state_for_muni_plot <- emergency_state %>%
  right_join(muni_to_state_map, by = "uf") %>%
  filter(!is.na(epi_week))
```

## Visualizing Risk Assessment and Emergency Declarations in 2024

```{r function-risk-plot}
create_risk_plot <- function(data, facet_var = NULL) {
  
  # Base plot
  p <- data %>%
    ggplot(aes(x = epi_week)) +
    # Incidence bars colored by risk level
    geom_col(aes(y = incidence_current, fill = riskassessment), width = 1, alpha = 0.8) +
    # Endemic channel 
    geom_ribbon(aes(ymin = low, ymax = high, fill = "Endemic Channel"), alpha = 0.5) +
    # Median incidence line
    geom_line(aes(y = median_incidence, color = "Historical Median", linetype = "Historical Median"), linewidth = 0.8) +
    # Scale x-axis
    scale_x_continuous(breaks = seq(min(data$epi_week), max(data$epi_week), by=5)) +
    # Define custom colors and labels for the fill aesthetic (Risk Levels)
    scale_fill_manual(
       name = "Risk Level & Baseline",
  values = c(
    "Endemic Channel" = "grey",
    "0" = "green",
    "1" = "yellow",
    "2" = "orange",
    "3" = "red"
  ),
  labels = c(
    "Endemic Channel" = "Endemic Channel",
    "0" = "Level 0 - Low Risk",
    "1" = "Level 1 - Initial Response",
    "2" = "Level 2 - Alert",
    "3" = "Level 3 - Emergency"), 
  drop = FALSE,
  limits = c("Endemic Channel", "0", "1", "2", "3")) +
  # Define manual color for historical Median line
  scale_color_manual(
  values = c("Historical Median" = "blue")) +
  # Define manual linetype for historical Median line
  scale_linetype_manual(
  values = c("Historical Median" = "dashed")) +
  # Combine fill and line legends
  guides(
    fill = guide_legend(title = NULL, order = 1),
    color = guide_legend(title = NULL, order = 2, override.aes = list(linetype = "dashed", color = "blue")),
    linetype = "none") +
theme_classic(base_size = 14) +
    theme(
      plot.title = element_text(face = "bold", size = rel(1.2)),
      axis.title = element_text(face = "bold", size = rel(1)),
      legend.position = "bottom",
      legend.title = element_text(face = "bold", size = rel(1)),
      legend.text = element_text(size = rel(0.9)),
      panel.grid.major.y = element_line(color = "grey90", linewidth = 0.5), # Add subtle grid lines
      strip.background = element_rect(fill = "grey92", color = NA), # Style facet titles
      strip.text = element_text(face = "bold", size = rel(1)),
      legend.box = "vertical"
    ) +
    labs(
      y = "Incidence Rate (per 100,000)", x = "Epidemiological Week")

  # Add faceting if a facet variable is provided
  if (!is.null(facet_var)) {
    p <- p + facet_wrap(vars(!!sym(facet_var)), scales = "free_y")
  }
  
  return(p)
}
```

### Country-wide

#### Dengue
```{r risk-country-dengue, fig.width=12, fig.height=10}
dengue_risk_country <- create_risk_plot(
  data = risk_2024_country %>% filter(virus == "Dengue")) +
  geom_vline(data = emergency_country, aes(xintercept = epi_week), linetype = "solid", color = "black", size = 1) +
  geom_vline(data = emergency_state, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1) +
  geom_text(data = emergency_state_labels, aes(x = epi_week, y = 220, label = uf_labels), vjust = 1.5, size = 3.5)

print(dengue_risk_country)
ggsave(here("Graphs/Risk Levels of Dengue Outbreak in 2024 in Brazil with Emergency Dec.png"), dengue_risk_country, height = 10, width = 12, dpi = 300, create.dir = TRUE)
```

#### Dengue, Chikungunya, and Zika
```{r risk-country-threee, fig.width=12, fig.height=10}
sum_risk_country <- create_risk_plot(
  data = risk_2024_country %>% filter(virus == "Sum of Arboviruses")) +
  geom_vline(data = emergency_country, aes(xintercept = epi_week), linetype = "solid", color = "black", size = 1) +
  geom_vline(data = emergency_state, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1) +
  geom_text(data = emergency_state_labels, aes(x = epi_week, y = 230, label = uf_labels), vjust = 1.5, size = 3.5)

print(sum_risk_country)
ggsave(here("Graphs/Risk Levels of Arbovirus (Sum) Outbreak in 2024 in Brazil with Emergency Dec.png"), sum_risk_country, height = 10, width = 12, dpi = 300, create.dir = TRUE)
```

### State-level 

#### Dengue
```{r risk-state-dengue, fig.width=12, fig.height=10}
dengue_risk_state <- create_risk_plot(
  data = risk_2024_state %>% filter(virus == "Dengue"),
  facet_var = "uf") + 
  geom_vline(data = emergency_state, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1.5)

print(dengue_risk_state)
ggsave(here("Graphs/Risk Levels of Dengue Outbreak in 2024 per State with Emergency Dec.png"), dengue_risk_state, height = 10, width = 12, dpi = 300, create.dir = TRUE)
```
    
#### Dengue, Chikungunya, and Zika
```{r risk-state-threee, fig.width=12, fig.height=10}
sum_risk_state <- create_risk_plot(
  data = risk_2024_state %>% filter(virus == "Sum of Arboviruses"),
  facet_var = "uf") + 
  geom_vline(data = emergency_state, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1.5)

print(sum_risk_state)
ggsave(here("Graphs/Risk Levels of Arbovirus (Sum) Outbreak in 2024 per State with Emergency Dec.png"), sum_risk_state, height = 10, width = 12, dpi = 300, create.dir = TRUE)
```

### Selected Municipalities

#### Dengue
```{r risk-muni-dengue, fig.width=12, fig.height=10}
dengue_risk_muni <- create_risk_plot(
  data = risk_2024_muni %>% filter(virus == "Dengue"),
  facet_var = "muni_name") +
  geom_vline(data = emergency_muni, aes(xintercept = epi_week), linetype = "dotted", color = "black", size = 1.5) +
  geom_vline(data = emergency_state_for_muni_plot, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1.5)

print(dengue_risk_muni)
ggsave(here("Graphs/Risk Levels of Dengue Outbreak in 2024 per Municipality with Emergency Dec.png"), dengue_risk_muni, height = 10, width = 12, dpi = 300, create.dir = TRUE)
```

#### Dengue, Chikungunya, and Zika
```{r risk-muni-threee, fig.width=12, fig.height=10}
sum_risk_muni <- create_risk_plot(
  data = risk_2024_muni %>% filter(virus == "Sum of Arboviruses"),
  facet_var = "muni_name") +
  geom_vline(data = emergency_muni, aes(xintercept = epi_week), linetype = "dotted", color = "black", size = 1.5) +
  geom_vline(data = emergency_state_for_muni_plot, aes(xintercept = epi_week), linetype = "dashed", color = "black", size = 1.5)

print(sum_risk_muni)
ggsave(here("Graphs/Risk Levels of Arbovirus (Sum) Outbreak in 2024 per Municipality with Emergency Dec.png"), sum_risk_muni, height = 10, width = 12, dpi = 300, create.dir = TRUE)
```

## Mapping First Week Alert in 2024

```{r mapping-firstweekalert, fig.width=12, fig.height=10}
# Filter for virus data and rename columns for joining
dengue_alert_week_state_filtered <- alert_week_state %>%
 filter(virus == "Dengue") %>%
 rename(uf = location_name)

dengue_alert_week_muni_filtered <- alert_week_muni %>%
 filter(virus == "Dengue") %>%
  rename(municipality = location_name)

# Join the alert week data with the geospatial data
# States
dengue_map_data_states <- states_sf %>%
 left_join(dengue_alert_week_state_filtered, by = c("abbrev_state" = "uf"))

# Capitals
dengue_map_data_capitals <- capitals_points %>%
 left_join(dengue_alert_week_muni_filtered, by = c("code_muni" = "ibge"))

# Breaks
dengue_week_breaks <- c(1, 5, 12, 53)

dengue_week_labels <- c("Week 1-4", "Week 5-12", "Week 13+")

dengue_map_data_states <- dengue_map_data_states %>%
 mutate(first_week_alert_2024 = as.numeric(first_week_alert_2024)) %>%
 mutate(alert_week_group = cut(first_week_alert_2024,
 breaks = dengue_week_breaks,
 labels = dengue_week_labels,
 include.lowest = TRUE,
 right = FALSE))
 
dengue_map_data_capitals <- dengue_map_data_capitals %>%
 mutate(first_week_alert_2024 = as.numeric(first_week_alert_2024)) %>%
 mutate(alert_week_group = cut(first_week_alert_2024,
 breaks = dengue_week_breaks,
 labels = dengue_week_labels,
 include.lowest = TRUE,
 right = FALSE))

# Generate the Map
#---------------------------------
dengue_risk_map_themed <- ggplot() +
 geom_sf(data = dengue_map_data_states, aes(fill = alert_week_group), color = "black", linewidth = 0.5) +
 geom_sf(data = dengue_map_data_capitals, aes(shape = alert_week_group),
 fill = NA, color = "darkgrey", alpha = 0.7, size = 4, stroke = 2) +
scale_fill_brewer(
    palette = "Greens",
    name = "First Week 'Alert' (States)",
 drop = TRUE
 ) +
 scale_shape_discrete(
 name = "First Week 'Alert' (Capitals)",
 drop = TRUE
 ) +
 theme_bw() +
 theme(
 legend.position = c(0.8, 0.1),
 legend.justification = c("left", "bottom"),
 legend.background = element_rect(fill = "transparent"),
 legend.box.background = element_blank(),
 panel.grid = element_blank(),
 panel.border = element_blank(),
 axis.title = element_text(),
 axis.text = element_text()
 )

print(dengue_risk_map_themed)
ggsave(here("Graphs", "Map First Week Alert 2024.png"), dengue_risk_map_themed, width = 12, height = 10, dpi = 300)
```