---
title: "Risk Assessment of Arboviruses in Brazilian State Capitals"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
lang: pt-BR
output:
  pdf_document:
    latex_engine: xelatex   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F)
library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(here)
library(haven)
library(lubridate)
library(readr)

loadfonts(device = "win")
```

# Content:

1.  Gathering Data

2.  Preparing Data

3.  Running Risk Assessment (for 2024, without epidemic years)

4.  Descriptive graphs of incidence and risk

5.  Control Diagrams (for 2024, without epidemic years)

6.  Risk Levels in 2024

# 1. Gathering Data

```{r get-data}
dengue_full <- read_csv(here("dengue/Dengue_Provaveis_2014_2024.csv")) %>% 
  mutate(epidemiologic_week = paste0(year, "w", week))
```

```{r}
#All arbovirsuses data:
arboviruses_full <- read.csv(here("data/cases_all_virus_2014_2024.csv"))


# Epidemiological week starting on the first week of the winter
epiweek40 <- read_dta(
  here("dengue/epidemiological_week_startsweek40.dta")) 

#List of municipalities of interest
codibge <- list(
  "Rio Branco"       = "120040",
  "Maceió"           = "270430",
  "Macapá"           = "160030",
  "Manaus"           = "130260",
  "Salvador"         = "292740",
  "Fortaleza"        = "230440",
  "Brasília"         = "530010",
  "Vitória"          = "320530",
  "Goiânia"          = "520870",
  "São Luís"         = "211130",
  "Cuiabá"           = "510340",
  "Campo Grande"     = "500270",
  "Belo Horizonte"   = "310620",
  "Belém"            = "150140",
  "João Pessoa"      = "250750",
  "Curitiba"         = "410690",
  "Recife"           = "261160",
  "Teresina"         = "221100",
  "Rio de Janeiro"   = "330455",
  "Natal"            = "240810",
  "Porto Alegre"     = "431490",
  "Porto Velho"      = "110020",
  "Boa Vista"        = "140010",
  "Florianópolis"    = "420540",
  "São Paulo"        = "355030",
  "Aracaju"          = "280030",
  "Palmas"           = "172100", 
  "Campinas"        =  "350950")

```

# 2. Preparing Data

## Selecting just the municipalities of interest

```{r}
arboviruses <- arboviruses_full %>% 
  filter(ibge %in% codibge)
```

## Grouping by municipality and epidemiological week

```{r prepare-data}
 ## Grouping by municipality and epidemiological week
arboviruses <- arboviruses %>% 
  # Correctly group by municipality, year, and week first
  group_by(ibge, class_regiao, municipality, year, week, Virus) %>% 
  summarise(
    cases = sum(likely_cases, na.rm = TRUE),
    population = sum(population, na.rm = TRUE), 
    deaths = sum(deaths, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    incidence = ifelse(population > 0, cases / (population / 100000), 0),
    wy = paste0(year, "w", as.integer(week))
  ) %>%
  full_join(epiweek40, by = "wy") %>%
  # IMPORTANT FIX: Group by municipality BEFORE arranging and filling NAs
  group_by(municipality) %>%
  arrange(year_week40, epidemiologic_week_startsweek40) %>% 
  filter(!is.na(municipality)) %>% 
  mutate(
    yw = as.character(paste0(year_week40,"-", epidemiologic_week_startsweek40)),
    epi_label = paste0(year_week40, "w", epidemiologic_week_startsweek40)
  ) %>%
  # Ungroup at the end to have a clean, final dataframe
  ungroup()
```

# 3. Risk Assessment

## Function for Risk Assessment

```{r risk-assess}
risk_analysis <- function(df, analysis_year = 2024) {
  # Defines a function that takes a dataframe 'df' and an 'analysis_year'.

  tryCatch({
    # If an error occurs, the code will jump to the 'error' section at the end.

    # === Part 1: Identify Epidemic Years based on Risk Score ===
    # The goal of this part is to analyze all historical data to determine which years behaved like epidemics.

    # Create a preliminary endemic channel using ALL historical data.
    preliminary_channel <- df %>%
      # 1. Select only data from years before the analysis_year.
      filter(year_week40 < analysis_year) %>%
      # 2. Group data by the epidemiological week number.
      group_by(week2 = epidemiologic_week_startsweek40) %>%
      # 3. For each week, calculate the historical average and standard deviation of incidence.
      summarise(avg_incidence = mean(incidence, na.rm = TRUE), sd_incidence = sd(incidence, na.rm = TRUE), .groups = "drop") %>%
      # 4. Calculate the upper alert threshold (95% confidence interval).
      mutate(high = avg_incidence + (1.96 * sd_incidence))

    # Check if there is enough historical data to proceed.
    if(nrow(preliminary_channel) == 0) return(NULL)

    # Assess all historical years against the preliminary channel to find epidemic years.
    risk_all_years <- df %>%
      # 1. Select only historical data.
      filter(year_week40 < analysis_year) %>%
      # 2. Add the calculated avg_incidence and high threshold to each row based on its week.
      left_join(preliminary_channel, by = c("epidemiologic_week_startsweek40" = "week2")) %>%
      # 3. Group the data by year to analyze each year independently.
      group_by(year_week40) %>%
      # 4. Ensure weeks are sorted chronologically within each year.
      arrange(epidemiologic_week_startsweek40, .by_group = TRUE) %>%
      # 5. Create several new columns (triggers) based on different risk conditions.
      mutate(
        # Trigger: Is incidence above average but below the high threshold? (1=yes, 0=no)
        increasedincidence = ifelse(incidence > avg_incidence & incidence < high, 1, 0),
        # Trigger: Has 'increasedincidence' been true for 4 straight weeks?
        increasedincidence4weeks = zoo::rollapply(increasedincidence, 4, function(x) as.integer(all(x == 1)), fill = NA, align = "right"),
        # Trigger: Is incidence above the high threshold?
        increasedincidencehigh = ifelse(incidence > high, 1, 0),
        # Trigger: Has 'increasedincidencehigh' been true for 4 straight weeks?
        increasedincidencehigh4weeks = zoo::rollapply(increasedincidencehigh, 4, function(x) as.integer(all(x == 1)), fill = NA, align = "right"),
        # Trigger: Did deaths increase from the previous week?
        increaseddeath = ifelse(deaths > lag(deaths), 1, 0),
        # Trigger: Has 'increaseddeath' been true for 4 straight weeks?
        increaseddeath4w = zoo::rollapply(increaseddeath, 4, function(x) as.integer(all(x == 1)), fill = NA, align = "right"),
        # Calculate log of incidence to check for exponential growth.
        ln_incidence = ifelse(incidence > 0, log(incidence), 0),
        # Calculate the week-over-week change in log(incidence).
        exponentialgrowth = ln_incidence - dplyr::lag(ln_incidence),
        # Trigger: Is growth exponential AND is incidence above the high threshold?
        increasedexpgrowth = ifelse(exponentialgrowth > 0 & incidence > high, 1, 0),
        # Trigger: Has 'increaseddeath' been true for 5 straight weeks?
        increaseddeath5w = zoo::rollapply(increaseddeath, 5, function(x) as.integer(all(x == 1)), fill = NA, align = "right")
      ) %>%
      # Replace any NAs created by the rolling functions (at the start of the series) with 0. #NÃO TENHO CERTEZA SE ESTA PARTE É NECESSÁRIA OU CORRETA
      #mutate(across(c(increasedincidence4weeks, increasedincidencehigh4weeks, increaseddeath4w, increasedexpgrowth, increaseddeath5w), ~replace_na(., 0))) %>%
      # Assign risk levels based on the triggers.
      mutate(
        # Risk Level 1: Sustained increase in cases.
        risk_level1 = if_else(increasedincidence4weeks == 1, 1L, 0L),
        # Risk Level 2: High alert for cases or sustained increase in deaths.
        risk_level2 = if_else(increasedincidencehigh4weeks == 1 | increaseddeath4w == 1, 2L, 0L),
        # Risk Level 3: Epidemic level based on exponential growth or prolonged increase in deaths.
        risk_level3 = if_else(increasedexpgrowth == 1 | increaseddeath5w == 1, 3L, 0L),
        # The final risk assessment for a week is the highest level triggered.
        riskassessment = pmax(risk_level1, risk_level2, risk_level3, na.rm = TRUE)
      )

    # From the historical assessment, create a list of years that hit risk level 3.
    epidemic_years <- risk_all_years %>% filter(riskassessment >= 3) %>% distinct(year_week40) %>% pull(year_week40)
    # Create a list of all available historical years.
    all_hist_years <- unique(df$year_week40[df$year_week40 < analysis_year])
    # Create the list of non-epidemic years by removing epidemic years from all historical years.
    non_epidemic_years <- all_hist_years[!all_hist_years %in% epidemic_years]
    # Select only the 5 most recent non-epidemic years for the final baseline.
    non_epidemic_years <- non_epidemic_years %>%
      sort() 

    # Fallback condition:  Skips the municipality if there are not enough non-epidemic years to build a reliable baseline (at least 2 are needed).
    if (length(non_epidemic_years) < 2) {
      
      message(paste("Skipping municipality:", 
                    unique(df$municipality), 
                    "| Reason: Fewer than 2 non-epidemic years found."))
                    
      return(NULL) 
    }

    # === Part 2: Build Final Control Diagram and Assess the Current Year ===
    # The goal of this part is to use the clean "non-epidemic" years to build a
    # final, reliable baseline and assess the current analysis_year against it.

    # Isolate the incidence and death data for the current analysis_year.
    incidence_current_year <- df %>% filter(year_week40 == analysis_year) %>% transmute(incidence_current = incidence, deaths, week2 = epidemiologic_week_startsweek40)

    # Build the final, clean endemic channel.
    control_data <- df %>%
      # 1. IMPORTANT: Filter for only the selected non-epidemic years.
      filter(year_week40 %in% non_epidemic_years) %>%
      # 2. Group by week.
      group_by(week2 = epidemiologic_week_startsweek40) %>%
      # 3. Recalculate the average and standard deviation using this cleaner data.
      summarise(avg_incidence = mean(incidence, na.rm = TRUE), sd_incidence = sd(incidence, na.rm = TRUE), .groups = "drop") %>%
      # 4. Create the final upper and lower thresholds for the control diagram.
      mutate(high = avg_incidence + 1.96*sd_incidence, low = pmax(0, avg_incidence - 1.96*sd_incidence)) %>%
      # 5. Join the final channel data with the current year's data.
      left_join(incidence_current_year, by = "week2")

    # Perform the final risk assessment for the analysis_year.
    risk_final <- control_data %>%
      arrange(week2) %>%
      mutate(
        increasedincidence = ifelse(incidence_current > avg_incidence & incidence_current < high, 1, 0),
        increasedincidence4weeks = zoo::rollapply(increasedincidence, 4, function(x) as.integer(all(x==1)), fill=NA, align="right"),
        increasedincidencehigh = ifelse(incidence_current > high, 1, 0),
        increasedincidencehigh4weeks = zoo::rollapply(increasedincidencehigh, 4, function(x)as.integer(all(x==1)), fill=NA, align="right"),
        increaseddeath = ifelse(deaths > lag(deaths), 1, 0),
        increaseddeath4w = zoo::rollapply(increaseddeath, 4, function(x) as.integer(all(x==1)), fill=NA, align="right"),
        ln_incidence = ifelse(incidence_current > 0, log(incidence_current), 0),
        exponentialgrowth = ln_incidence - dplyr::lag(ln_incidence),
        increasedexpgrowth = ifelse(exponentialgrowth > 0 & incidence_current > high, 1, 0),
        increaseddeath5w = zoo::rollapply(increaseddeath, 5, function(x) as.integer(all(x==1)), fill=NA, align="right")
      ) %>%
      mutate(across(c(increasedincidence4weeks, increasedincidencehigh4weeks, increaseddeath4w, increasedexpgrowth, increaseddeath5w), ~replace_na(., 0))) %>%
      mutate(
        risk_level1 = if_else(increasedincidence4weeks == 1, 1L, 0L),
        risk_level2 = if_else(increasedincidencehigh4weeks == 1 | increaseddeath4w == 1, 2L, 0L),
        risk_level3 = if_else(increasedexpgrowth == 1 | increaseddeath5w == 1, 3L, 0L),
        riskassessment = pmax(risk_level1, risk_level2, risk_level3, na.rm = TRUE)
      )

    # Return the results as a list containing two objects.
    return(list(risk_final = risk_final, epidemic_years = epidemic_years))

  }, error = function(e) {
    # If any error occurred in the 'try' block, this code will run.
    message(paste("Could not process municipality:", unique(df$municipality), "| Error:", e$message))
    return(NULL)
  })
}
```

## Running risk assessment

```{r risks-listed}
# Nest the data, creating a list-column with a dataframe for each municipality
nested_data <- arboviruses %>%
  group_by(municipality, Virus) %>%
  nest()

# Apply function to each nested dataframe
results <- nested_data %>%
  mutate(risk_analysis_results = map(data, risk_analysis))

# Unnest the results for the 2024 risk assessment
risk_2024 <- results %>%
  # Safely extract the 'risk_final' tibble from the list
  mutate(risk_data = map(risk_analysis_results, ~ .x$risk_final)) %>%
  select(municipality, Virus, risk_data) %>%
  unnest(cols = c(risk_data))
```

## Table of epidemic years per municipality

```{r epidemic-years-per-muni}
# Helper function to safely extract epidemic years from the results list
extract_epidemic_years <- function(model_output) {
  if (!is.null(model_output) && "epidemic_years" %in% names(model_output)) {
    if (length(model_output$epidemic_years) > 0) {
      return(tibble(epidemic_year = as.character(model_output$epidemic_years)))
    }
  }
  return(tibble(epidemic_year = character(0)))
}

# Extract and unnest the epidemic years data
epidemic_years_data <- results %>%
  mutate(epidemic_years = map(risk_analysis_results, extract_epidemic_years)) %>%
  select(municipality, epidemic_years) %>%
  unnest(cols = c(epidemic_years))

# Create a summary table
epidemic_summary_table <- epidemic_years_data %>%
  group_by(municipality) %>%
  summarise(
    epidemic_years_list = paste(sort(unique(epidemic_year)), collapse = ", "),
    .groups = "drop"
  ) %>%
  rename(
    `Municipality` = municipality,
    `Epidemic Years` = epidemic_years_list
  )

# Table
knitr::kable(epidemic_summary_table, caption = "Epidemic Years of Dengue by Municipality")
```

# 4. Descriptive graphs of incidence and risk

## Plotting Incidence per Municipality per Epidemiological Week

```{r plot-suspected}
#Dengue:
arboviruses %>% 
  filter(Virus == "Dengue") %>% 
ggplot(aes(x = yw, y = incidence, group = 1)) +
  geom_line(color = "steelblue") +
  facet_wrap(~municipality, scales = "free_y") + 
  labs(
    title = "Weekly Dengue Incidence by Municipality",
    x = "Epidemiological Week (Year-Week)",
    y = "Incidence Rate (per 100,000 people)"
  ) +
  scale_x_discrete(breaks = levels(arboviruses$yw)[seq(1, length(levels(arboviruses$yw)))]) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8))
```

```{r}
#Zika:
arboviruses %>% 
  filter(Virus == "Zika") %>% 
ggplot(aes(x = yw, y = incidence, group = 1)) +
  geom_line(color = "steelblue") +
  facet_wrap(~municipality, scales = "free_y") + 
  labs(
    title = "Weekly Zika Incidence by Municipality",
    x = "Epidemiological Week (Year-Week)",
    y = "Incidence Rate (per 100,000 people)"
  ) +
  scale_x_discrete(breaks = levels(arboviruses$yw)[seq(1, length(levels(arboviruses$yw)))]) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8))
```

```{r}
#Chikungunya:
arboviruses %>% 
  filter(Virus == "Chikungunya") %>% 
ggplot(aes(x = yw, y = incidence, group = 1)) +
  geom_line(color = "steelblue") +
  facet_wrap(~municipality, scales = "free_y") + 
  labs(
    title = "Weekly Chikungunya Incidence by Municipality",
    x = "Epidemiological Week (Year-Week)",
    y = "Incidence Rate (per 100,000 people)"
  ) +
  scale_x_discrete(breaks = levels(arboviruses$yw)[seq(1, length(levels(arboviruses$yw)))]) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8))
```

## Plotting Risk per Municipality per Epidemiological Week

```{r risk-per-city}
#Dengue
risk_2024 %>% 
  filter(Virus == "Dengue") %>% 
  ggplot(aes(x = week2, y = riskassessment, group = 1)) +
  # geom_step is great for visualizing discrete changes in risk level
  geom_step(color = "firebrick") + 
  # Create a separate plot for each municipality
  facet_wrap(~ municipality) + 
  labs(
    title = "Dengue Risk Assessment for 2024 by Municipality",
    x = "Epidemiological Week of 2024",
    y = "Calculated Risk Level"
  ) +
  scale_y_continuous(breaks = 0:3, limits = c(0, 3)) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 8))
```

# 5. Control Diagrams (for 2024, without epidemic years)

```{r plot-risk-2024}
risk_2024 %>% 
  filter(Virus == "Dengue") %>% 
ggplot(aes(x = week2)) +
  
  # The historical endemic channel (grey ribbon)
  geom_ribbon(aes(ymin = low, ymax = high), fill = "lightblue", alpha = 0.8) +
  
  # The historical average line
  geom_line(aes(y = avg_incidence, linetype = "Historical Average", group = 1), color = "black") +
  
  # The 2024 incidence line
  geom_line(aes(y = incidence_current, linetype = "2024 Incidence", group = 1), color = "red") +
  
  # Create a grid of plots for each municipality
  facet_wrap(~ municipality, scales = "free_y") +
  
  # Customize the legend for clarity
  scale_linetype_manual(
    name = "Legend",
    values = c("Historical Average" = "dashed", "2024 Incidence" = "solid")
  ) +
  
  # Add titles and labels
  labs(
    title = "Control Diagrams for Dengue Incidence (2024)",
    subtitle = "2024 Incidence vs. Non-Epidemic Historical Average",
    y = "Incidence Rate (per 100,000)",
    x = "Epidemiological Week of 2024"
  ) +
  theme_minimal() +
  theme(legend.position = "top")
```

# 6. Risk Levels in 2024

```{r}
#Preparing the data for creating the graphs
risk_2024 <- risk_2024 %>%  
  filter(!is.na(municipality)) %>%
  mutate(week2 = as.numeric(week2), 
         riskassessment = factor(riskassessment))

every_10 <- seq(min(risk_2024$week2, na.rm = TRUE), max(risk_2024$week2, na.rm = TRUE), by = 10)
plot_labels_10 <- risk_2024$week2[match(every_10, risk_2024$week2)]

```

### Dengue

```{r, warning=FALSE}

#Plot
g1 <- risk_2024 %>% 
  filter(Virus == "Dengue") %>% 
  ggplot(aes(x = week2)) +
  geom_rect(aes(xmin = week2, xmax = week2 + 1,
              ymin = low, ymax = high),
          fill = "darkgrey")+
  geom_rect(aes(xmin = week2, xmax = week2 + 1,
                ymin = 0, ymax = incidence_current,
                fill = riskassessment)) +
  geom_line(aes(y = avg_incidence,
                color = "Average incidence (prev. 5 years)",
                linetype = "Average incidence (prev. 5 years)"),
            size = 1) +
  geom_step(aes(y = incidence_current,
                color = "black",
                linetype = "solid"),
            size = 1) +
  theme_bw() +  
  scale_color_manual(name = "Historical Average Incidence",
                     values = c(
    "Average incidence (prev. 5 years)" = "blue"
  )) +
  scale_linetype_manual(name = "", values = c(
    "Average incidence (prev. 5 years)" = "dashed"
  )) +
scale_fill_manual( values = c("0" = "#07E000", "1" = "#D9E000", "2" = "#E07000", "3" = "#E10000", "Endemic Channel Upper Limit" = "darkgrey"), name = "Risk Level & Baseline", 
                    labels = c("0" = "Level 0 - Normal", 
               "1" = "Level 1 - Alert", 
               "2" = "Level 2 - High",
               "3" = "Level 3 - Epidemic",
               "Endemic Channel Upper Limit" = "Endemic Channel Upper Limit")) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "bottom",
    text = element_text(family = "Times New Roman"))+
    scale_x_continuous(
    breaks = every_10,
    labels = plot_labels_10) +
  # Add titles and labels
  labs(
    title = "Risk Levels of Dengue Outbreak in 2024",
    subtitle = "Weekly incidence compared to the historical endemic channel (excluding epidemic years)",
    y = "Incidence Coefficient (per 100,000 inhabitants)",
    x = "Epidemiological Week",
    caption = "Data from: DATASUS. Analysis and Visualization by the authors")+
  facet_wrap(~municipality, scales = "free_y", ncol = 4)

g1

ggsave(here("Graphs/Risk Levels of Dengue Outbreak in 2024.png"),g1, height = 15, width = 15)
```

### Zika

```{r, warning=FALSE}

#Plot
g1 <- risk_2024 %>% 
  filter(Virus == "Zika") %>% 
  ggplot(aes(x = week2)) +
  geom_rect(aes(xmin = week2, xmax = week2 + 1,
              ymin = low, ymax = high),
          fill = "darkgrey")+
  geom_rect(aes(xmin = week2, xmax = week2 + 1,
                ymin = 0, ymax = incidence_current,
                fill = riskassessment)) +
  geom_line(aes(y = avg_incidence,
                color = "Average incidence (prev. 5 years)",
                linetype = "Average incidence (prev. 5 years)"),
            size = 1) +
   geom_step(aes(y = incidence_current,
                color = "black",
                linetype = "solid"),
            size = 1) +
  theme_bw() +  
  scale_color_manual(name = "Historical Average Incidence",
                     values = c(
    "Average incidence (prev. 5 years)" = "blue"
  )) +
  scale_linetype_manual(name = "", values = c(
    "Weekly Incidence" = "solid",
    "Average incidence (prev. 5 years)" = "dashed"
  )) +
scale_fill_manual( values = c("0" = "#07E000", "1" = "#D9E000", "2" = "#E07000", "3" = "#E10000", "Endemic Channel Upper Limit" = "darkgrey"), name = "Risk Level & Baseline", 
                    labels = c("0" = "Level 0 - Normal", 
               "1" = "Level 1 - Alert", 
               "2" = "Level 2 - High",
               "3" = "Level 3 - Epidemic",
               "Endemic Channel Upper Limit" = "Endemic Channel Upper Limit")) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "bottom",
    text = element_text(family = "Times New Roman"))+
    scale_x_continuous(
    breaks = every_10,
    labels = plot_labels_10) +
  # Add titles and labels
  labs(
    title = "Risk Levels of Zika Outbreak in 2024",
    subtitle = "Weekly incidence compared to the historical endemic channel (excluding epidemic years)",
    y = "Incidence Coefficient (per 100,000 inhabitants)",
    x = "Epidemiological Week",
    caption = "Data from: DATASUS. Analysis and Visualization by the authors")+
  facet_wrap(~municipality, scales = "free_y", ncol = 4)

g1

ggsave(here("Graphs/Risk Levels of Zika Outbreak in 2024.png"),g1, height = 15, width = 15)
```

### Chikungunya

```{r, warning=FALSE}

#Plot
g1 <- risk_2024 %>% 
  filter(Virus == "Chikungunya") %>% 
  ggplot(aes(x = week2)) +
  geom_rect(aes(xmin = week2, xmax = week2 + 1,
              ymin = low, ymax = high),
          fill = "darkgrey")+
  geom_rect(aes(xmin = week2, xmax = week2 + 1,
                ymin = 0, ymax = incidence_current,
                fill = riskassessment)) +
  geom_line(aes(y = avg_incidence,
                color = "Average incidence (prev. 5 years)",
                linetype = "Average incidence (prev. 5 years)"),
            size = 1) +
  geom_step(aes(y = incidence_current,
                color = "black", 
                linetype = "Incidence coefficient"),
            size = 1) +
  theme_bw() +  
  scale_color_manual(name = "Historical Average Incidence",
                     values = c(
    "Weekly Incidence" = "black",
    "Average incidence (prev. 5 years)" = "blue"
  )) +
  scale_linetype_manual(name = "", values = c(
    "Weekly Incidence" = "solid",
    "Average incidence (prev. 5 years)" = "dashed"
  )) +
scale_fill_manual( values = c("0" = "#07E000", "1" = "#D9E000", "2" = "#E07000", "3" = "#E10000", "Endemic Channel Upper Limit" = "darkgrey"), name = "Risk Level & Baseline", 
                    labels = c("0" = "Level 0 - Normal", 
               "1" = "Level 1 - Alert", 
               "2" = "Level 2 - High",
               "3" = "Level 3 - Epidemic",
               "Endemic Channel Upper Limit" = "Endemic Channel Upper Limit")) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "bottom",
    text = element_text(family = "Times New Roman"))+
    scale_x_continuous(
    breaks = every_10,
    labels = plot_labels_10) +
  # Add titles and labels
  labs(
    title = "Risk Levels of Chikungunya Outbreak in 2024",
    subtitle = "Weekly incidence compared to the historical endemic channel (excluding epidemic years)",
    y = "Incidence Coefficient (per 100,000 inhabitants)",
    x = "Epidemiological Week",
    caption = "Data from: DATASUS. Analysis and Visualization by the authors")+
  facet_wrap(~municipality, scales = "free_y", ncol = 4)

g1

ggsave(here("Graphs/Risk Levels of Chikungunya Outbreak in 2024.png"),g1, height = 15, width = 15)
```
