---
title: "Risk Assessment of Arboviruses in Brazilian State Capitals - 2024"
author: "Luiz C Soares"
date: "2025-08-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Steps for replication:

1.  Gathering the data

2.  Preparing data for evaluation

3.  Creating Control Diagrams (for all years)

4.  Making Risk Assessment (for all years)

5.  Creating Control Diagrams (for 2024, without epidemic years)

6.  Making Risk Assessment (for 2024, without epidemic years)

7.  Creating descriptive graphs of incidence and risk

```{r}


```

#First version

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(zoo)
library(ggplot2)
library(here)
library(haven)
library(lubridate)

```

# 2. Preparing data for evaluation

```{r load-data}
dengue <- read_csv(here("dengue/Dengue_Provaveis_2014_2024.csv"))

dengue <- dengue %>% 
  mutate(epidemiologic_week = paste0(year, "w", week))
```

# Campinas Data Preparation

```{r campinas-prep}
campinas <- dengue %>% 
  filter(municipality == "350950 CAMPINAS") %>% # Selected city: Campinas/SP
  group_by(epidemiologic_week, week, year) %>% 
  summarise( #* Cases, deaths, and pop for this period
    cases = sum(likely_cases, na.rm=TRUE),
    population = sum(population, na.rm=TRUE),
    deaths = sum(deaths, na.rm=TRUE),
    .groups="drop"
  ) %>%
  mutate( #* Dengue incidence
    incidence = ifelse(population > 0, cases / (population/100000), 0), 
    wy = paste0(year, "w", as.integer(week)), #* Week to variable
    epi_label = paste0(year, "w", week)
  )
```

### Merging with epiweek

```{r}

# Epidemiological week starting on the first week of the winter
epiweek40 <- read_dta(
  here("dengue/epidemiological_week_startsweek40.dta")) 

campinas <- campinas %>%
  left_join(epiweek40, by = "wy")%>% 
  mutate(yw = as.character(paste0(year_week40,"-", epidemiologic_week_startsweek40))) %>% 
  filter(!is.na(incidence),
         !is.na(week40))

campinas <- campinas %>%
  mutate(yw = factor(yw, levels = unique(yw)))  # keep the original order

ggplot(campinas, aes(x = yw, y = incidence, group = 1)) +
  geom_line() +
  labs(x = "Epidemiological Week", 
       y = "Incidence coefficient (Suspected Cases)") +
  scale_x_discrete(breaks = levels(campinas$yw)[seq(1, length(levels(campinas$yw)), by = 25)]) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

# Control diagram (average + SD without epidemic years)

### Incidence for 2024

```{r incidence-2024}
incidence2024 <- campinas %>% 
  filter(year == 2024) %>%
  transmute(incidence2024 = incidence, deaths, week2 = week, wy)
```

```{r control-diagram}
control_data <- campinas %>%
  filter(year %in% c(2016,2017,2018,2020,2021)) %>%
  group_by(week2 = week) %>%
  summarise(
    avg_incidence = mean(incidence, na.rm=TRUE),
    sd_incidence = sd(incidence, na.rm=TRUE),
    .groups="drop"
  ) %>%
  mutate(high = avg_incidence + 1.96*sd_incidence,
         low = pmax(0, avg_incidence - 1.96*sd_incidence)) %>%
  left_join(incidence2024, by="week2")

# Plot endemic channel
ggplot(control_data, aes(x=week2, group = 1)) +
  geom_ribbon(aes(ymin=low, ymax=high), fill="grey80") +
  geom_line(aes(y=avg_incidence), linetype="dashed", color="black") +
  geom_line(aes(y=incidence2024), color="red") +
  labs(y="Incidence coefficient", x="Epidemiological week")
```

# Risk Assessment for 2024

```{r risk-assessment}
risk <- control_data %>%
  arrange(week2) %>%
  mutate(
    riskassessment = 0,
    increasedincidence = ifelse(incidence2024 > avg_incidence & incidence2024 < high, 1, 0),
    increasedincidence4weeks = zoo::rollapply(increasedincidence, 4, function(x) all(x==1), fill=NA, align="right"),
    increasedincidencehigh = ifelse(incidence2024 > high, 1, 0),
    increasedincidencehigh4weeks = zoo::rollapply(increasedincidencehigh, 4, function(x) all(x==1), fill=NA, align="right"),
    increaseddeath = c(NA, diff(deaths) > 0),
    increaseddeath4w = zoo::rollapply(increaseddeath, 4, function(x) all(x==1), fill=NA, align="right"),
    ln_incidence = log(incidence2024),
    exponentialgrowth = ln_incidence - dplyr::lag(ln_incidence),
    increasedexpgrowth = ifelse(exponentialgrowth > 0 & incidence2024 > high, 1, 0),
    increaseddeath5w = zoo::rollapply(increaseddeath, 5, function(x) all(x==1), fill=NA, align="right")
  ) %>%
  mutate(
    riskassessment = case_when(
      increasedincidence4weeks == 1 ~ 1,
      increasedincidencehigh4weeks == 1 | increaseddeath4w == 1 ~ 2,
      increasedexpgrowth == 1 | increaseddeath5w == 1 ~ 3,
      TRUE ~ 0
    )
  )

saveRDS(risk, here("dengue/Campinas/riskCampinas_2024.RDS"))

ggplot(risk, aes(x=week2, y=riskassessment, group = 1)) +
  geom_line() +
  scale_y_continuous(breaks=0:3) +
  labs(x="Epidemiological week", y="Risk Assessment")
```
